---
title: '#1: Isotope data standardization'
author: 
- Nadege Belouard^[UMR EcoBio, UMR DECOD nadege.belouard@gmail.com]
date: "03/06/2024"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

The purpose of this first vignette is to complete the stable isotope data transformation from raw data to standardized data to be used in stable isotope niche calculations. Data standardization will allow comparison between different ecosystems.

First, we load the packages that we are going to use for this vignette.
```{r setup, warning = F, message = F}
library(dplyr)
library(ggplot2)
library(knitr)
library(tidyr)
library(magrittr)
library(here)
```

We then load the raw dataset and have a look at it.
```{r load data}
Dataset <- read.csv(file.path(here(), "data_raw", "Raw_data.csv"), header=T)
Dataset %>% head() %>% kable()
```

The dataset contains the pond code, sample taxa (species x life stage), species and group (amphibians, crayfish, baseline consumers), sample ID and weight, and finally stable isotope values in the columns "D15N" and "D13C", with the C/N ratio in the last column.


We generate a few useful tables:
Detail of the number of samples per taxa per pond
```{r number of samples}
table(Dataset$Pond, Dataset$Taxa)
```


Calculate the mean sample weight per taxa
```{r sample weights}
Dataset %>%
  group_by(Group) %>%
  summarise(meanweight = round(mean(Weight*1000),0), 
            sdweight = round(sd(Weight*1000),0)) %>% 
  kable()
```


# A- Data wrangling

## 1- Correction of D13C for samples with high lipid content (C/N > 3.5)

First, let's have a look at the average C/N for each species.

```{r look at C.N ratio, message=FALSE, warning=FALSE}
Dataset %>% 
  group_by(Taxa) %>%
  summarize(meanCN = mean(C.N),
            minCN = min(C.N),
            maxCN = max(C.N)) %>%
  kable()
```

Since many of them have C/N higher than 3.5, we need to correct the D13C for high lipid content for all groups. We use the equation of Post, 2007. The lipid-free D13C is saved in a new column named "D13C_LF".

```{r correct D13C for high lipid content using Post equation}

Dataset %<>% mutate(D13C_LF = D13C - 3.32 + 0.99 * C.N)
```

And we check that the correction increased D13C by comparing it to y = x:

```{r plot the D13C correction, fig.height=4, fig.width=6}
plot(Dataset$D13C_LF ~ Dataset$D13C,
     xlab = "original D13C",
     ylab = "corrected D13C") +
  abline(0,1)
```

Corrected D13C are above the y = x line, so it looks ok.


## 2- Correction of amphibian fin values for muscle values

In accordance with institutional ethics, we sampled amphibian fins instead of muscle, and need to correct values according to equations published in Bélouard et al. (2018);

The equations used to correct the $D^{13}C$ values are:  
For the agile frog: $D^{13}C_{muscle} = 0.85 \times D^{13}C_{fin} - 5.50$  
For the European tree frog:	$D^{13}C_{muscle} = 0.87 \times D^{13}C_{fin} - 4.80$  
For the marbled newt: $D^{13}C_{muscle} = 0.70 \times D^{13}C_{fin} - 7.43$  
For the palmate newt:	$D^{13}C_{muscle} = 0.71 \times D^{13}C_{fin} - 7.03$  
  
The equations used to correct the $D^{15}N$ values are:  
For the agile frog: $D^{15}N_{muscle} = 0.86 \times D^{15}N_{fin} + 0.97$  
For the European tree frog: $D^{15}N_{muscle} = 1.04 \times D^{15}N_{fin} + 0.67$  
For the marbled newt: $D^{15}N_{muscle} = 1.01 \times D^{15}N_{fin} - 0.48$  
For the palmate newt: $D^{15}N_{muscle} = 0.89 \times D^{15}N_{fin} + 0.40$  
  
Values corrected for this translation are stored in two new columns: D13C_LFMu and D15N_Mu.

```{r correct fin values}

for (i in 1:dim(Dataset)[1]) {
  if (Dataset$Taxa[i] == "Agile frog") {
    Dataset$D13C_LFMu[i] = 0.85 * Dataset$D13C_LF[i] -5.50
    Dataset$D15N_Mu[i] = 0.86 * Dataset$D15N[i] + 0.97
  } else if (Dataset$Taxa[i] == "Tree frog") {
    Dataset$D13C_LFMu[i] = 0.87 * Dataset$D13C_LF[i] -4.80
    Dataset$D15N_Mu[i] = 1.04 * Dataset$D15N[i] + 0.67
  } else if (Dataset$Taxa[i] == "Marbled newt") {
    Dataset$D13C_LFMu[i] = 0.70 * Dataset$D13C_LF[i] - 7.43
    Dataset$D15N_Mu[i] = 1.01 * Dataset$D15N[i] - 0.48
  } else if (Dataset$Taxa[i] == "Palmate newt") {
    Dataset$D13C_LFMu[i] = 0.71 * Dataset$D13C_LF[i] - 7.03
    Dataset$D15N_Mu[i] = 0.89 * Dataset$D15N[i] + 0.40
  } else { 
    Dataset$D13C_LFMu[i] = Dataset$D13C_LF[i]
    Dataset$D15N_Mu[i] = Dataset$D15N[i]
  }
}
```





# B - Data standardization

We need primary consumers to serve as a baseline to standardize the isotopic values. Candidate taxa are: clams (Corbiculidae), Physa, Gammarus,  Corixidae, Asellus, and zooplankton. Here are the sample sizes for each pond.

``` {r candidate primary consumers for each pond, message = F, warning = F}

x <- Dataset %>%
  dplyr::filter(Taxa %in% c("Corbiculidae", "Physa", "Zooplankton", "Gammarus",
                            "Corixidae", "Asellus")) %>%
  group_by(Taxa, Pond)%>%
  summarise(n = n()) %>%
  spread(Pond, n)

kable(x)
```

## 1- Trophic position

Standardized D15N values are called "trophic position" (TP). We standardize N values using strict primary consumers: clams and Physa. We calculate the average D15N of the primary consumers and add it to the dataset.

``` {r calculate the baseline D15N, message = F, warning = F}
MeanBaselineN <- Dataset %>%
  dplyr::filter(Taxa %in% c("Corbiculidae", "Physa")) %>%
  group_by(Pond, Taxa)%>%
  summarise(meanD15N = mean(D15N_Mu)) %>%
  group_by(Pond) %>%
  summarise(meanD15N = mean(meanD15N))

# Add mean baseline to each individual of Dataset
Dataset <- merge(Dataset, MeanBaselineN, by = "Pond", all = T)
```

Verify that we still have the correct number of rows in the table
```{r check nb rows 1}
dim(Dataset)[1] == 1306
```

Based on this baseline of primary consumers, we calculate the trophic position (TP) as  
$TP = 2 + D15N_{sample} - D15N_{baseline} / delta$
Delta is the fractionation coefficient and here delta = 3.4

``` {r calculate TP}
Delta = 3.4
Dataset %<>% mutate(TP = 2+(Dataset$D15N_Mu-Dataset$meanD15N)/Delta)
```


## 2- Carbon correction

We need taxa to serve as markers of the variability in carbon resources in each pond, to standardize the C isotopic value of samples into a variable called "Delta 13C corrected". Taxa used are: clams, physa, zooplankton, gammarus, corixidae and asellus. We compute the mean of their averaged values ($D^{13}C_{meanC1}$), then correct individual carbon values (i) using the equation:   
$D^{13}C_{cor} = (D^{13}C_{i} - D^{13}C_{meanC1}) / range D^{13}C_{C1}$


``` {r carbon correction by primary consumers, message = F, warning = F}
BaselineC <- Dataset %>%
  filter(Taxa %in% c("Corbiculidae", "Physa", "Zooplankton", "Gammarus", 
                     "Corixidae", "Asellus")) %>%
  group_by(Pond, Taxa) %>%
  summarize(meanD13C = mean(D13C_LFMu)) %>%
  group_by(Pond) %>%
  summarize(meanavD13C = mean(meanD13C),
            maxavD13C = max(meanD13C),
            minavD13C = min(meanD13C))

#Add mean baseline to each individual of Dataset
Dataset <- merge(Dataset, BaselineC, by = "Pond", all = TRUE)

#Calcul Ccor
Dataset %<>% mutate(D13Ccor_consoI = (D13C_LFMu - meanavD13C) / 
                      (maxavD13C - minavD13C))
```


We now have standardized isotope values!

# C- Export dataset

Verify that we still have the correct number of rows and keep only the taxa of interest for the next steps.
```{r check nb rows 2}
dim(Dataset)[1] == 1306

Dataset %<>% filter(Group %in% c("Amphibians", "Crayfish"))
```

Export the dataset
``` {r export dataset}

# Create a folder if it does not exist
if (file.exists(file.path(here(),"exported_data")) == FALSE){
  dir.create(file.path(here(),"exported_data"))
}

write.csv(Dataset, file.path(here(), "exported_data", 
                             "Isotope_data_standardized.csv"), row.names = F)
```

In the next vignettes, we will install the package siar, then compute the niche metrics necessary for the analysis of the stable isotope niches.