---
title: "#2: Compute niche metrics"
author: 
- Nadege Belouard^[UMR EcoBio, UMR DECOD, nadege.belouard@gmail.com]
date: "03/06/2024"
output:
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

This vignette computes all the population stable isotope niche metrics.
We first load the packages.

```{r setup, warning = F, message = F}
library(SIBER)
library(dplyr)
library(ggplot2)
library(here)
library(magrittr)
library(tidyr)
library(tibble)
library(knitr)
library(reshape2)
library(spatstat.utils)
```

We then load the dataset and keep only the variables that will be useful. 
```{r load dataset}

Dataset <- read.csv(file.path(here(), "exported_data", 
                                       "Isotope_data_standardized.csv"), 
                    header=T)

Dataset %<>% dplyr::select(Pond, Taxa, Species, Group, Sample_ID, 
                    Weight, TP, D13Ccor_consoI)
  
head(Dataset)
```


# 1- Crayfish ontogenetic shift

First we are looking into whether there is an ontogenetic shift in crayfish. 
Wee load the crayfish length data
```{r load dataset}

# load file
CrayfishLength <- read.table(file.path(here(), "data_raw", 
                                       "CrayfishLength.txt"), 
                    header=T)

# keep only the stable isotope data for crayfish
Crayfish <- Dataset %<>% filter(Species == "Crayfish")

# join the two datasets to obtain the length and stable isotope values for each individual
Crayfish %<>% left_join(CrayfishLength, by = "Sample_ID")
# check that all individuals have been associated with a length
dim(Crayfish %>% filter(is.na(Length)))[1] == 0
```

Plot the data to see if there is an ontogenetic shift
```{r graph shift}

# Isotopic biplot
ggplot(data = Crayfish, aes(x = TP, y = D13Ccor_consoI)) +
  geom_point(aes(col = Taxa)) +
  facet_wrap(~Pond)
``` 

Test the ontogenetic shift on the carbon values first
```{r test shift on carbon values}
# Plot the D13C data against the crayfish length
ggplot(data = Crayfish, aes(y = D13Ccor_consoI, x = Length)) +
  geom_point(aes(), show.legend = F) +
  geom_smooth(method = lm, se = FALSE, col = "darkgray") +
  facet_wrap(~Pond) +
  xlab("Crayfish length (mm)") +
  ylab(bquote(''*delta^13~C["cor"]*'')) +
  theme_classic()
# There is an ncrease in D13Ccor with length, except in P06, P10, P14, P16

# Run the linear models for each pond
Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P05")) 
summary(Onto13C)
# P05 D13Ccor_consoI = -0.944143 + 0.006091 * length, p = 1.68e-06 ***
shapiro.test(Onto13C$residuals)

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P07")) 
summary(Onto13C)
# P07 y = -0.4014810 + 0.0051519 x, p = 7.19e-12 ***

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P08")) 
summary(Onto13C)
# P08 y = -0.1880660 + 0.006687 x, p = 7.36e-13 ***

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P09")) 
summary(Onto13C)
# P09 y = -0.2867501 + 0.0047571 x, p = 7.82e-16 ***

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P10")) 
summary(Onto13C)
# P10 y = 0.127039 - 0.003261 x, p = 0.00865 ** NEGATIVE ESTIMATE

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P13")) 
summary(Onto13C)
# P13 y = -0.1429704 + 0.0041330 x, p = 1e-08 ***

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P14")) 
summary(Onto13C)
# P14 y = 0.0562290 -0.0012756 x, p = 0.132 NS and NEGATIVE ESTIMATE

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P15")) 
summary(Onto13C)
# P15 y = 0.0181422 + 0.0021203 x, p = 5.78e-05 ***

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P16")) 
summary(Onto13C)
# P16 y = 0.257888 - 0.002912 x, p = 0.0237 * NEGATIVE ESTIMATE

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P17")) 
summary(Onto13C)
# P17 y = -0.7509301 + 0.0101905 x, p = 1.14e-11 ***

Onto13C <- lm(D13Ccor_consoI ~ Length, data = Crayfish %>% filter(Pond == "P18")) 
summary(Onto13C)
# P18 y = -0.238652 + 0.006569 x, p = 7.66e-08 ***
``` 

Now test for nitrogen values
```{r test shift on nitrogen values} 

#Plot the ontogenetic shift on nitrogen values against crayfish length
ggplot(data = Crayfish, aes(x = Length, y = TP)) +
  geom_point(aes(), show.legend = F) +
  geom_smooth(method = lm, se = FALSE, col = "darkgray") +
  facet_wrap(~Pond) +
  xlab("Crayfish length (mm)") +
  # ylab("TP") +
  theme_classic()

# Run the linear model for each pond
OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P05"))
summary(OntoTP)
shapiro.test(OntoTP$residuals)
# P05 y = 2.6661974 -0.0035418 x, p = 5.7e-05 ***, NEGATIVE ESTIMATE

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P07"))
summary(OntoTP)
# P07 y = 2.864e+00 + 2.975e-05 x, p = 0.961, NS

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P08"))
summary(OntoTP)
# P08 y = 2.350218 + 0.011484 x, p = 1.83e-12 ***, 

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P09"))
summary(OntoTP)
# P09 y = 2.2342960+ 0.0046668x, p =7.23e-08 ***, 

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P10"))
summary(OntoTP)
# P10 y = 2.5093372+ 0.0039351x, p =8.31e-08 ***, 

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P13"))
summary(OntoTP)
# P13 y = 2.346057-0.005903 x, p =0.000108 ***,  NEGATIVE ESTIMATE

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P14"))
summary(OntoTP)
# P14 y = 2.4414659+ 0.0019453x, p =0.0156 *

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P15"))
summary(OntoTP)
# P15 y = 2.3161938+ 0.0009667x, p =0.113, NS

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P16"))
summary(OntoTP)
# P16 y = 2.510389+ 0.002087x, p =0.187, NS

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P17"))
summary(OntoTP)
# P17 y = 2.196241+ 0.002532x, p =0.105, NS

OntoTP <- lm(TP ~ Length, data = Crayfish %>% filter(Pond == "P18"))
summary(OntoTP)
# P18 y = 1.8632440+ 0.0039885x, p =0.000158 ***
```

# 2- Population metrics

We calculate population metrics for amphibians and crayfish.

We begin with selecting the species of interest and removing populations with inadequate samples sizes, i.e. large populations with n < 10. Some populations have n < 10 but are kept because we have good indicators from CMR that the sample size is representative of the population.

```{r select appropriate populations, message = F, warning = F}

# We want to exclude target populations with n < 10, except three pops: marbled newt in ponds P04 and P11, and palmate newt in pop P10 where we assume we have sampled the entire population.
Dataset %>% group_by(Species, Pond) %>% summarise(n = n()) %>% 
  filter(n < 10)

Dataset<-Dataset[!Dataset$Species=="Crayfish"|
                                 !Dataset$Pond == "P06",]
Dataset<-Dataset[!Dataset$Species=="Agile frog"|
                                 !Dataset$Pond=="P02",]
Dataset<-Dataset[!Dataset$Species=="Agile frog"|
                                 !Dataset$Pond=="P17",]
Dataset<-Dataset[!Dataset$Species=="Marbled newt"|
                                 !Dataset$Pond=="P08",]
Dataset<-Dataset[!Dataset$Species=="Marbled newt"|
                                 !Dataset$Pond=="P10",]
Dataset<-Dataset[!Dataset$Species=="Palmate newt"|
                                 !Dataset$Pond=="P07",]
Dataset<-Dataset[!Dataset$Species=="Palmate newt"|
                                 !Dataset$Pond== "P13",]
Dataset<-Dataset[!Dataset$Species=="Palmate newt"|
                                 !Dataset$Pond== "P14",]
Dataset<-Dataset[!Dataset$Species=="Palmate newt"|
                                 !Dataset$Pond=="P16",]
Dataset<-Dataset[!Dataset$Species=="Palmate newt"|
                                 !Dataset$Pond=="P17",]


# Separate per niche (species x pond)
spx <- split(Dataset$D13Ccor_consoI, 
             list(Dataset$Pond, Dataset$Species), drop = TRUE)
spy <- split(Dataset$TP, 
             list(Dataset$Pond, Dataset$Species), drop = TRUE)



# # In the case where we consider separately adult and juvenile crayfish (for a figure in Supp. Mat), run the following lines too
# Dataset %>% filter(Species == "Crayfish") %>% group_by(Taxa, Pond) %>% summarise(n = n()) %>% 
#   filter(n < 10)
# 
# Dataset<-Dataset[!Dataset$Taxa=="Juv crayfish"|
#                                  !Dataset$Pond=="P09",]
# Dataset<-Dataset[!Dataset$Taxa=="Juv crayfish"|
#                                  !Dataset$Pond=="P16",]
# Dataset<-Dataset[!Dataset$Taxa=="Juv crayfish"|
#                                  !Dataset$Pond=="P17",]
# 
# spx <- split(Dataset$D13Ccor_consoI, 
#              list(Dataset$Pond, Dataset$Taxa), drop = TRUE)
# spy <- split(Dataset$TP, 
#              list(Dataset$Pond, Dataset$Taxa), drop = TRUE)
```

Check if the number of individuals is correct
```{r check number of individuals is correct}

dim(Dataset[Dataset$Taxa=="Agile frog",])[1] == 254 
dim(Dataset[Dataset$Taxa=="Tree frog",])[1] == 168
dim(Dataset[Dataset$Taxa=="Palmate newt",])[1] == 192
dim(Dataset[Dataset$Taxa=="Marbled newt",])[1] == 101
dim(Dataset[Dataset$Species=="Crayfish",])[1] == 348

```


## a- SEA coordinates 

We then generate the coordinates of the ellipses so we can plot them later.

```{r population SEAc coordinates, warning = F, message = F}

Pop <- NULL
coordY <- NULL
coordX <- NULL
j = 1

for (i in 1:length(spx)) {
  SE <- standard.ellipse(spx[[i]],spy[[i]],steps=1)
  Pop[j:(j+360)]<-names(spx)[i]
  coordY[j:(j+360)]<-SE$ySEAc[1:361]
  coordX[j:(j+360)]<-SE$xSEAc[1:361]
  j=j+361
}

CoordSEAc <- cbind(Pop,coordY,coordX)
summary(CoordSEAc)
CoordSEAc <- as.data.frame(CoordSEAc)
CoordSEAc$coordY <- as.numeric(as.character(CoordSEAc$coordY))
CoordSEAc$coordX <- as.numeric(as.character(CoordSEAc$coordX))
str(CoordSEAc)
CoordSEAc$Pond <- gsub("\\..*","",CoordSEAc$Pop)
CoordSEAc$Species <- gsub(".*\\.","",CoordSEAc$Pop)
head(CoordSEAc)
CoordSEAc$Pond <- as.factor(CoordSEAc$Pond)
levels(CoordSEAc$Pond)

write.csv(CoordSEAc, 
          file.path(here(), "exported_data", "Coordinates_SEAc_populations.csv"), 
          row.names = F)

## For the table corresponding to the ellipses separated by juvenile and adult crayfish, save it in the following file:
# write.csv(CoordSEAc, 
#           file.path(here(), "exported_data", "Coordinates_SEAc_ontogeneticshift.csv"), 
#           row.names = F)
``` 


Calculate the mean position of standard ellipses (centroids) at the pond level
```{r centroides of standard ellipses}

Mean_position <- CoordSEAc %>% group_by(Pond, Species, Pop) %>%
  summarise(TPavg = mean(coordY),
            D13Cavg = mean(coordX))

write.csv(Mean_position, file.path(here(), "exported_data", 
                                   "Meanposition_SEAc.csv"), row.names = F)

```




## b- Ellipse areas and overlap at the pond scale

Calculate the area and percentage of ellipse overlap at the population level

```{r area of ellipses overlap: initialize vectors}

#Split dataset by species
Crayfish <- Dataset %>% filter(Species == "Crayfish")
Treefrog <- Dataset %>% filter(Species == "Tree frog")
Agilefrog <- Dataset %>% filter(Species == "Agile frog") 
Palmatenewt <- Dataset %>% filter(Species == "Palmate newt")
Marblednewt <- Dataset %>% filter(Species == "Marbled newt") 


#For each species, split dataset by pond
spx_Agilefrog <- split(Agilefrog$D13Ccor_consoI, list(Agilefrog$Pond), 
                       drop = FALSE)
spy_Agilefrog <- split(Agilefrog$TP, list(Agilefrog$Pond), 
                       drop = FALSE)
spx_Treefrog <- split(Treefrog$D13Ccor_consoI, list(Treefrog$Pond), 
                       drop = FALSE)
spy_Treefrog <- split(Treefrog$TP, list(Treefrog$Pond), 
                       drop = FALSE)
spx_Palmatenewt <- split(Palmatenewt$D13Ccor_consoI, list(Palmatenewt$Pond), 
                       drop = FALSE)
spy_Palmatenewt <- split(Palmatenewt$TP, list(Palmatenewt$Pond), 
                       drop = FALSE)
spx_Marblednewt <- split(Marblednewt$D13Ccor_consoI, list(Marblednewt$Pond), 
                       drop = FALSE)
spy_Marblednewt <- split(Marblednewt$TP, list(Marblednewt$Pond), 
                       drop = FALSE)
spx_Crayfish <- split(Crayfish$D13Ccor_consoI, list(Crayfish$Pond), 
                       drop = FALSE)
spy_Crayfish <- split(Crayfish$TP, list(Crayfish$Pond), 
                       drop = FALSE)

#Initialize vectors
Pond = NULL
Species1 = NULL
Species2 = NULL
AreaOverlap = NULL
AreaSp1 = NULL
AreaSp2 = NULL
Table = data.frame()
```

Loop over all crayfish-amphibian pairs
```{r loop over all crayfish-amphibian pairs}

for (i in names(spx_Crayfish)) {
  if (length(spx_Agilefrog[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Crayfish", after = length(Species1))
    Species2 <- append(Species2, "Agile frog", after = length(Species2))
    Overlap <- siar::overlap(spx_Crayfish[[i]], spy_Crayfish[[i]],
                             spx_Agilefrog[[i]], spy_Agilefrog[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Treefrog[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Crayfish", after = length(Species1))
    Species2 <- append(Species2, "Tree frog", after = length(Species2))
    Overlap <- siar::overlap(spx_Crayfish[[i]], spy_Crayfish[[i]],
                             spx_Treefrog[[i]], spy_Treefrog[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Palmatenewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Crayfish", after = length(Species1))
    Species2 <- append(Species2, "Palmate newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Crayfish[[i]], spy_Crayfish[[i]],
                             spx_Palmatenewt[[i]], spy_Palmatenewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Marblednewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Crayfish", after = length(Species1))
    Species2 <- append(Species2, "Marbled newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Crayfish[[i]], spy_Crayfish[[i]],
                             spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
}
```

Loop between amphibian species pairs
```{r area of ellipses overlap: loop between amphibians}

for (i in names(spx_Agilefrog)) {
  if (length(spx_Treefrog[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Agile frog", after = length(Species1))
    Species2 <- append(Species2, "Tree frog", after = length(Species2))
    Overlap <- siar::overlap(spx_Agilefrog[[i]], spy_Agilefrog[[i]],
                             spx_Treefrog[[i]], spy_Treefrog[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Palmatenewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Agile frog", after = length(Species1))
    Species2 <- append(Species2, "Palmate newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Agilefrog[[i]], spy_Agilefrog[[i]],
                             spx_Palmatenewt[[i]], spy_Palmatenewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Marblednewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Agile frog", after = length(Species1))
    Species2 <- append(Species2, "Marbled newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Agilefrog[[i]], spy_Agilefrog[[i]],
                             spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
}



for (i in names(spx_Treefrog)) {
  if (length(spx_Palmatenewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Tree frog", after = length(Species1))
    Species2 <- append(Species2, "Palmate newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Treefrog[[i]], spy_Treefrog[[i]],
                             spx_Palmatenewt[[i]], spy_Palmatenewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Marblednewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Tree frog", after = length(Species1))
    Species2 <- append(Species2, "Marbled newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Treefrog[[i]], spy_Treefrog[[i]],
                             spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
}


for (i in names(spx_Palmatenewt)) {
  if (length(spx_Marblednewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Palmate newt", after = length(Species1))
    Species2 <- append(Species2, "Marbled newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Palmatenewt[[i]], spy_Palmatenewt[[i]],
                             spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
}

```

Merge results in a table and export it
``` {r analyses of population overlaps}

Table <- cbind(Pond, Species1, Species2, AreaOverlap = round(AreaOverlap,4), 
             AreaSp1 = round(AreaSp1,4), AreaSp2 = round(AreaSp2,4))

Table <- as.data.frame(Table)
Table$Prct_overlap <- round(as.numeric(Table$AreaOverlap) / (as.numeric(Table$AreaSp1)+as.numeric(Table$AreaSp2)),4)
Table$Prct_overlap_sp2 <- round(as.numeric(Table$AreaOverlap) / as.numeric(Table$AreaSp2),4)
Table$Prct_overlap_sp1 <- round(as.numeric(Table$AreaOverlap) / as.numeric(Table$AreaSp1),4)

head(Table)

write.csv(Table, file.path(here(), "exported_data", "Overlap_populations.csv"), 
          row.names = F)

```

Finally, generate a table with all SEAc at the pond scale
```{r calculate population ellipses}

Dataset_ellipses <- Dataset %>% 
  dplyr::select("iso1" = D13Ccor_consoI, "iso2" = TP, "group" = Species, 
                "community" = Pond) %>%
  as.data.frame()

Dataset_SIBER <- createSiberObject(Dataset_ellipses)
Pop_metrics_SEAc <- groupMetricsML(Dataset_SIBER)

#Table transformation
Pop_metrics_ellipses <- as.data.frame(t(Pop_metrics_SEAc))
Pop_metrics_ellipses$Code <- rownames(Pop_metrics_ellipses)
Pop_metrics_ellipses$Pond <- gsub("[.].*","",Pop_metrics_ellipses$Code)
Pop_metrics_ellipses$Species <- gsub(".*.[.]","",Pop_metrics_ellipses$Code)

write.csv(Pop_metrics_ellipses[-1], 
          file.path(here(), "exported_data", "Population_metrics.csv"), 
          row.names = F)
```


# Long table
Generate the long version of this table combining SEAc and mean position for Supplementary Material

```{r generate the long version of this table for Supp Mat}

Pop_metrics <- read.csv(file.path(here(), "exported_data", 
                                  "Population_metrics.csv"))

Mean_position <- read.csv(file.path(here(), "exported_data", 
                                    "Meanposition_SEAc.csv"))

names(Mean_position)[3] = "Code"
Mean_position %<>% dplyr::select(-Pond, -Species) 
Pop_metrics <- merge(Pop_metrics, Mean_position, by = "Code")
Pop_metrics %<>% dplyr::select(Pond, Species, SEAc, TPavg, D13Cavg)

Pop_metrics_long <- rbind(
      Pop_metrics %>% dcast(Pond ~ Species, value.var = "SEAc") %>% add_column(Metrics = "SEAc"),
      Pop_metrics %>% dcast(Pond ~ Species, value.var = "TPavg") %>% add_column(Metrics = "TPavg"),
      Pop_metrics %>% dcast(Pond ~ Species, value.var = "D13Cavg") %>% add_column(Metrics = "D13Cavg"))
      

write.csv(Pop_metrics_long, file.path(here(), "exported_data", "Population_metrics_long.csv"), row.names = F)
```

The next step will be to plot these communities.