---
title: "Local isotopic niche partitioning and the coexistence between native amphibians and invasive crayfish #4: Network plots with ellipses"
author: "Nadege Belouard"
date: "23/09/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

This vignette draws all the plots related to the trophic networks. First, we load the different packages and datasets that we need.

```{r setup}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(magrittr)
library(here)

# Points
Dataset <- read.csv(file.path(here(), "exported_data", 
                              "Isotope_data_standardized.csv"), header=T)

# Ellipses
#Per population
CoordSEAc_pop <- read.csv(file.path(here(), 
                                    "exported_data", 
                                    "Coordinates_SEAc_populations.csv"), 
                          header=T)
names(CoordSEAc_pop)[1] <- "Taxa"

#Per species
CoordSEAc_sp <- read.csv(file.path(here(), "exported_data", 
                                   "Coordinates_SEAc_regional.csv"), header=T)


# Average points
Avg_data_complete <- Dataset %>% group_by(Taxa, Pond) %>%
  summarise(TPavg = mean(TP),
            D13Cavg = mean(D13Ccor_consoI),
            TPsd = sd(TP),
            D13Csd = sd(D13Ccor_consoI))


```

We summarize each taxa by the average point and its SD to graphically represent the position and the variability in the isotopic signature of each taxa.


# Plots with ellipses per species

## Figure 1
```{r plot network with ellipses per species and all individuals, include=FALSE}

# With outliers
M_ind <- Dataset %>% filter (Group %in% c("Crayfish", "Amphibians"))
M_coord <- CoordSEAc_sp
M_ind$Species <- factor(M_ind$Species, levels = c("Crayfish",
                                                    "Agile frog",
                                                    "Tree frog",
                                                    "Palmate newt",
                                                    "Marbled newt"))

species_ellipses <- ggplot(data = M_ind, aes(x = D13Ccor_consoI, y = TP)) +
  geom_point(aes(col = Species), alpha = 0.7, shape = 16) +
  geom_path(data = M_coord, aes(coordX, coordY, col = Species), lwd = 1.5) +
  scale_color_manual(values = c("#E69F00", 
                                "#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab(bquote(''*delta^13~C["cor"]*'')) +
  guides(col = guide_legend("Species")) +
  ylab("Trophic position") +
  theme_classic() + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))

ggsave(species_ellipses, filename = file.path(here(), "figures", 
                                              "ellipses_species.jpg"), 
       height = 4, width = 6)


# figure with ellipses only
species_ellipses <- ggplot() +
  geom_path(data = M_coord, aes(coordX, coordY, col = Species), lwd = 1.5) +
  scale_color_manual(values = c("#0072B2", "#E69F00",
                                          "#CC79A7", "#56B4E9", 
                                          "#009E73")) +
  xlab("D13Ccor") +
  ylab("Trophic position") +
  guides(col = guide_legend("Species")) +
  theme_classic()

ggsave(species_ellipses, filename = file.path(here(), "figures", "ellipses_species_wopoints.jpg"), height = 4, width = 6)

```



# Plots with ellipses per population per pond

## Figure 2
```{r plot ellipses only, include=FALSE}

M_ind$Species <- factor(M_ind$Species, levels = c("Crayfish",
                                                    "Agile frog",
                                                    "Tree frog",
                                                    "Palmate newt",
                                                    "Marbled newt"))

# Facet wrap ellipses
ellipses_pop <- ggplot() +
  theme_classic() +
  geom_path(data = CoordSEAc_pop, aes(coordX, coordY, col = Species), 
            size=2, linetype=1) +
  scale_color_manual(values = c("#0072B2", "#E69F00",
                                          "#CC79A7", "#56B4E9", 
                                          "#009E73")) +
  facet_wrap(~Pond, scales = "free_x") +
  xlab(bquote(''*delta^13~C["cor"]*'')) +
  ylab("Trophic position")

ellipses_pop
ggsave(ellipses_pop, filename = file.path(here(), "figures", 
                                          "ellipses_pop.jpg"), 
       width = 15, height = 10)


# Set limits for x values
blank_data <- data.frame(Pond_code = c("U01", "0U1", 
                                   "U02", "U02", 
                                   "U03", "U03",
                                   "U04", "U04",
                                   "U05", "U05",
                                   "U06", "U06",
                                   "I01", "I01",
                                   "I02", "I02",
                                   "I03", "I03",
                                   "I04", "I04",
                                   "I05", "I05",
                                   "I06", "I06",
                                   "I07", "I07",
                                   "I08", "I08",
                                   "I09", "I09",
                                   "I10", "I10",
                                   "I11", "I11",
                                   "I12", "I12",
                                   "I13", "I13",
                                   "I14", "I14"),
                         y = 1,
                         x = c(-1.5, 1.5, 
                               -2, 1.5, 
                               -4, 1.5,
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5))
blank_data$Pond_code <- factor(blank_data$Pond_code, 
                               levels = c("U1", "U2", "U3", "U4", "U5", "U6",
                                          "I1", "I2", "I3", "I4", "I5", "I6",
                                          "I7", "I8", "I9", "I10", "I11", "I12",
                                          "I13", "I14"))

# Facet wrap ellipses and points
# Dataset_target <- Dataset %>% filter(!TaxaUsed == "Other")

ellipses_pop <- ggplot(data = M_ind, aes(D13Ccor_consoI, TP)) +
  theme_classic() +
  geom_point(aes(col = Species), alpha = 0.7) +
  geom_path(data = CoordSEAc_pop, aes(coordX, coordY, col = Species), 
            size=1, linetype=1) +
  scale_color_manual(values = c("#E69F00", 
                                "#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  geom_blank(data = blank_data, aes(x = x, y = y)) +
  facet_wrap(~Pond, scales = "free_x") +
  xlab(bquote(''*delta^13~C["cor"]*'')) +
  ylab("Trophic position") +
  guides(col = guide_legend("Species"))

ellipses_pop

ggsave(ellipses_pop, filename = file.path(here(), "figures",
                                          "ellipses_pop_points.jpg"),
       width = 15, height = 10)

```
