---
title: "Stable coexistence between four native and one invasive species seen through the lens of isotopic niche analyses #4: Compute niche metrics"
author: "Nadege Belouard"
date: "11/20/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This vignette computes all isotopic niche metrics relative to populations and communities.
We first load the packages and the dataset, and check if the dataset has the correct number of lines.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(viridis)
library(siar)
library(SIBER)
library(dplyr)
library(spatstat.utils)
library(ggplot2)
library(here)
library(magrittr)
library(tidyr)
library(tibble)
library(reshape2)
```

```{r load datasets}

# Full dataset, with C1, crayfish and amphibians
Dataset_complete <- read.csv(file.path(here(), "exported_data", "Isotope_data_standardized_reduced.csv"), header=T)

# Same dataset without 3 outliers
# Dataset <- read.csv(file.path(here(), "exported_data", "Isotope_data_minusoutliers.csv"), header=T)

# dim(Dataset)[1] == 1443
dim(Dataset_complete)[1] == 1446

# Generate the variables "TaxaUsed" and "Group"
for (i in 1:length(Dataset_complete$Pond)) {
  if (Dataset_complete$Taxa[i] %in% c("Agile frog", "Tree frog", "Palmate newt", "Marbled newt")) {
    Dataset_complete$TaxaUsed[i] = Dataset_complete$Taxa[i]
  } else if(Dataset_complete$Taxa[i] %in% c("Ad crayfish", "Juv crayfish")) {
    Dataset_complete$TaxaUsed[i] = "Crayfish"
  } else{
    Dataset_complete$TaxaUsed[i] = "Other"
  }
}

# for (i in 1:length(Dataset$Pond)) {
#   if (Dataset$Taxa[i] %in% c("Agile frog", "Tree frog", "Palmate newt", "Marbled newt")) {
#     Dataset$TaxaUsed[i] = Dataset$Taxa[i]
#   } else if(Dataset$Taxa[i] %in% c("Ad crayfish", "Juv crayfish")) {
#     Dataset$TaxaUsed[i] = "Crayfish"
#   } else{
#     Dataset$TaxaUsed[i] = "Other"
#   }
# }

# Weight per sample group
# Dataset %>% filter(!Pond %in% c("J", "K", "88")) %>% 
#   group_by(Group) %>% 
#   summarise(meanweight = round(mean(Weight*1000),0), sdweight = round(sd(Weight*1000),0))

#Number of samples?
# Nbsamples <- table(Dataset$Pond, Dataset$Species_precis)

```



# 1- Regional niche (per species over all ponds)

## Dataset without outliers
```{r select appropriate populations, include=FALSE}

Subset_Dataset <- Dataset %>% dplyr::filter(!Group == "C1")
dim(Subset_Dataset)[1]
unique(Subset_Dataset$Pond)

Subset_Dataset %>% group_by(TaxaUsed) %>% summarise(meanTP = mean(TP),
                                                    meanD13C = mean(D13Ccor_consoI))

```

Layman metrics
```{r calculate Layman metrics for populations, include=FALSE}

# 6 niches (2 for crayfish, 4 for amphibians)
spx <- split(Subset_Dataset$D13Ccor_consoI, Subset_Dataset$Species, drop = TRUE)
length(spx)

spy <- split(Subset_Dataset$TP, Subset_Dataset$Species, drop = TRUE)
length(spy)

Table <- data.frame(metrics = c("dX_range", "dY_range", "NND", "SDNND", "CD", "TA"))

for (i in 1:length(spx)) {  
  Metrics <- laymanMetrics(spx[[i]],spy[[i]])
  Data <- as.data.frame(Metrics$metrics)
  Data$metrics <- rownames(Data)
  colnames(Data)[1] <- names(spx[i])
  Table <- merge(Table, Data, by = 'metrics', suffixes = colnames(Data)[1], all=TRUE)
}

Pop_metrics_Layman <- as.data.frame(t(Table))
colnames(Pop_metrics_Layman) <- Pop_metrics_Layman[1,]
Pop_metrics_Layman <- Pop_metrics_Layman[-1,]
Pop_metrics_Layman$Code <- rownames(Pop_metrics_Layman)

```

SEA metrics
```{r calculate population ellipses, include=FALSE}

Subset_Dataset_ellipses <- Subset_Dataset %>% 
  dplyr::select("iso1" = D13Ccor_consoI, "iso2" = TP, "group" = Species) %>%
  mutate(community = 1) %>% 
  as.data.frame()

Subset_Dataset_SIBER <- createSiberObject(Subset_Dataset_ellipses)
Subset_metrics_SEAc <- groupMetricsML(Subset_Dataset_SIBER)

#Table transformation
Subset_metrics_ellipses <- as.data.frame(t(Subset_metrics_SEAc))
Subset_metrics_ellipses$Code <- rownames(Subset_metrics_ellipses)
Subset_metrics_ellipses$Code <- gsub(".*.[.]","",Subset_metrics_ellipses$Code)

```

Merge datasets
```{r merge datasets, include=FALSE}

Subset_metrics <- merge(Pop_metrics_Layman, Subset_metrics_ellipses[-1], by = "Code")

write.csv(Subset_metrics, file.path(here(), "exported_data", "Subset_metrics_species.csv"), row.names = F)
```

Ellipse coordinates
```{r SEAc, include=FALSE}

Species<-NULL
coordY<-NULL
coordX<-NULL
j=1

for (i in 1:length(spx)) {
  SE <- standard.ellipse(spx[[i]],spy[[i]],steps=1)
  Species[j:(j+360)]<-names(spx)[i]
  coordY[j:(j+360)]<-SE$ySEAc[1:361]
  coordX[j:(j+360)]<-SE$xSEAc[1:361]
  j=j+361
}

CoordSEAc<-cbind(Species,coordY,coordX)
summary(CoordSEAc)
CoordSEAc<-as.data.frame(CoordSEAc)
CoordSEAc$coordY<-as.numeric(as.character(CoordSEAc$coordY))
CoordSEAc$coordX<-as.numeric(as.character(CoordSEAc$coordX))
str(CoordSEAc)
head(CoordSEAc)

write.csv(CoordSEAc, file.path(here(), "exported_data", "Coordinates_SEAc_species_subset.csv"), row.names = F)
``` 

Area of ellipses overlap
```{r area of ellipses overlap: initialize vectors}

#Initialize vectors
Species1 = NULL
Species2 = NULL
AreaOverlap = NULL
AreaSp1 = NULL
AreaSp2 = NULL
Table = data.frame()

#For each species, split dataset 
spx <- split(Subset_Dataset$D13Ccor_consoI, list(Subset_Dataset$TaxaUsed),drop=FALSE)
spy <- split(Subset_Dataset$TP, list(Subset_Dataset$TaxaUsed),drop=FALSE)

# Run command: with crayfish
for (i in names(spx[c(1,3:5)])){
Overlap <- siar::overlap(spx$Crayfish, spy$Crayfish,
                         spx[[i]], spy[[i]], steps=1)
AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
Species1 <- append(Species1, "Crayfish", after = length(Species1))
Species2 <- append(Species2, names(spx[i]), after = length(Species2))
AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
} 
 
# Run command: with agile frog
for (i in names(spx[c(3:5)])){
Overlap <- siar::overlap(spx$`Agile frog`, spy$`Agile frog`,
                         spx[[i]], spy[[i]], steps=1)
AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
Species1 <- append(Species1, "Agile frog", after = length(Species1))
Species2 <- append(Species2, names(spx[i]), after = length(Species2))
AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
} 

# Run command: with marbled newt
for (i in names(spx[c(4:5)])){
Overlap <- siar::overlap(spx$`Marbled newt`, spy$`Marbled newt`,
                         spx[[i]], spy[[i]], steps=1)
AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
Species1 <- append(Species1, "Marbled newt", after = length(Species1))
Species2 <- append(Species2, names(spx[i]), after = length(Species2))
AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
}


# Run command: with palmate newt/tree frog
Overlap <- siar::overlap(spx$`Palmate newt`, spy$`Palmate newt`,
                         spx$`Tree frog`, spy$`Tree frog`, steps=1)
AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
Species1 <- append(Species1, "Palmate newt", after = length(Species1))
Species2 <- append(Species2, "Tree frog", after = length(Species2))
AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))

```


Analysis of overlap areas
``` {r analyses of overlaps}

Table<-cbind(Species1, Species2, AreaOverlap = round(AreaOverlap,4), 
             AreaSp1 = round(AreaSp1,4), AreaSp2 = round(AreaSp2,4))

Table <- as.data.frame(Table)
Table$Prct_overlap <- round(as.numeric(Table$AreaOverlap) / (as.numeric(Table$AreaSp1)+as.numeric(Table$AreaSp2)),4)
Table$Prct_overlap_sp2 <- round(as.numeric(Table$AreaOverlap) / as.numeric(Table$AreaSp2),4)
Table$Prct_overlap_sp1 <- round(as.numeric(Table$AreaOverlap) / as.numeric(Table$AreaSp1),4)
Table

write.csv(Table, file.path(here(), "exported_data", "Overlap_species_subset.csv"), row.names = F)
```




## Complete dataset (incl. outliers)
```{r select appropriate species, include=FALSE}

Complete_Dataset <- Dataset_complete %>% dplyr::filter(!Group == "C1")
dim(Complete_Dataset)[1]

Complete_Dataset %>% group_by(TaxaUsed) %>% summarise(meanTP = mean(TP),
                                                      meanD13C = mean(D13Ccor_consoI))

```

Layman metrics
```{r calculate Layman metrics for species, include=FALSE}

# 6 niches (2 for crayfish, 4 for amphibians)
spx <- split(Complete_Dataset$D13Ccor_consoI, Complete_Dataset$Species, drop = TRUE)
length(spx)

spy <- split(Complete_Dataset$TP, Complete_Dataset$Species, drop = TRUE)
length(spy)

Table <- data.frame(metrics = c("dX_range", "dY_range", "NND", "SDNND", "CD", "TA"))

for (i in 1:length(spx)) {  
  Metrics <- laymanMetrics(spx[[i]],spy[[i]])
  Data <- as.data.frame(Metrics$metrics)
  Data$metrics <- rownames(Data)
  colnames(Data)[1] <- names(spx[i])
  Table <- merge(Table, Data, by = 'metrics', suffixes = colnames(Data)[1], all=TRUE)
}

Pop_metrics_Layman <- as.data.frame(t(Table))
colnames(Pop_metrics_Layman) <- Pop_metrics_Layman[1,]
Pop_metrics_Layman <- Pop_metrics_Layman[-1,]
Pop_metrics_Layman$Code <- rownames(Pop_metrics_Layman)

```


Now we calculate the equivalent using the bootstrapping method. For each pop, we sample 10 individuals in the population, calculate the 6 metrics of Layman, then do it again 999 times.

```{r calculate Layman metrics with bootstrapping for populations, include=FALSE}

# Make sure we find the same results everytime by setting the seed
set.seed(10)

# We group samples by pond and species again
spx <- split(Complete_Dataset$D13Ccor_consoI, 
             list(Complete_Dataset$TaxaUsed), drop = TRUE)
length(spx)

spy <- split(Complete_Dataset$TP, list(Complete_Dataset$TaxaUsed), drop = TRUE)
length(spy)

# Number of bootstrapped required
nr <- 9999 # the number of reps
ncol = 5

# Create empty vectors to store the results 
dNR <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
dCR <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
TA <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
CD <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
MNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
SDNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))

names(dNR)<-names(spx)
names(dCR)<-names(spx)
names(TA)<-names(spx)
names(CD)<-names(spx)
names(MNND)<-names(spx)
names(SDNND)<-names(spx)

for (j in 1:length(spx)) {
  if (length(spx[[j]]) > 10) {
    for (i in 1:nr) {
      simX<-sample(spx[[j]], 10, replace = FALSE, prob = NULL)
      simY<-sample(spy[[j]],10,replace = FALSE,prob=NULL)
      layman <- laymanmetrics(simX, simY)
      dNR[i,j] <- layman$dN_range
      dCR[i,j] <- layman$dC_range
      TA[i,j] <- layman$hull$TA
      CD[i,j] <- layman$CD
      MNND[i,j] <- layman$MNND
      SDNND[i,j] <- layman$SDNND
    }
    }
  }

```


```{r calculate average bootstrapped Layman metrics for populations, include=FALSE}

Pop_metrics_bootstrapped <- as.data.frame(cbind(NRb = colMeans(dNR), 
                                  CRb = colMeans(dCR), 
                                  TAb = colMeans(TA),
                                  CDb = colMeans(CD), 
                                  MNNDb = colMeans(MNND), 
                                  SDNNDb = colMeans(SDNND),
                                  Code = colnames(dNR)))

```



SEA metrics
```{r calculate population ellipses, include=FALSE}

Complete_Dataset_ellipses <- Complete_Dataset %>% 
  dplyr::select("iso1" = D13Ccor_consoI, "iso2" = TP, "group" = Species) %>%
  mutate(community = 1) %>% 
  as.data.frame()

Complete_dataset_SIBER <- createSiberObject(Complete_Dataset_ellipses)
Complete_metrics_SEAc <- groupMetricsML(Complete_dataset_SIBER)

#Table transformation
Complete_metrics_ellipses <- as.data.frame(t(Complete_metrics_SEAc))
Complete_metrics_ellipses$Code <- rownames(Complete_metrics_ellipses)
Complete_metrics_ellipses$Code <- gsub(".*.[.]","",Complete_metrics_ellipses$Code)

```


Merge datasets
```{r merge datasets, include=FALSE}

Pop_metrics <- merge(Pop_metrics_Layman, Pop_metrics_bootstrapped, by = "Code")
dim(Pop_metrics)

Pop_metrics <- merge(Pop_metrics, Complete_metrics_ellipses[-1], by = "Code")


write.csv(Pop_metrics, file.path(here(), "exported_data", "Complete_metrics_species.csv"), row.names = F)

```


Ellipse coordinates
```{r SEAc, include=FALSE}

# We group samples by pond and species again
spx <- split(Complete_Dataset$D13Ccor_consoI, 
             list(Complete_Dataset$TaxaUsed), drop = TRUE)
length(spx)

spy <- split(Complete_Dataset$TP, list(Complete_Dataset$TaxaUsed), drop = TRUE)
length(spy)

Species<-NULL
coordY<-NULL
coordX<-NULL
j=1

for (i in 1:length(spx)) {
  SE <- standard.ellipse(spx[[i]],spy[[i]],steps=1)
  Species[j:(j+360)]<-names(spx)[i]
  coordY[j:(j+360)]<-SE$ySEAc[1:361]
  coordX[j:(j+360)]<-SE$xSEAc[1:361]
  j=j+361
}

CoordSEAc<-cbind(Species,coordY,coordX)
summary(CoordSEAc)
CoordSEAc<-as.data.frame(CoordSEAc)
CoordSEAc$coordY<-as.numeric(as.character(CoordSEAc$coordY))
CoordSEAc$coordX<-as.numeric(as.character(CoordSEAc$coordX))
str(CoordSEAc)
head(CoordSEAc)

write.csv(CoordSEAc, file.path(here(), "exported_data", "Coordinates_SEAc_species_complete.csv"), row.names = F)
``` 

Area of ellipses overlap
```{r area of ellipses overlap: initialize vectors}

#Initialize vectors
Species1 = NULL
Species2 = NULL
AreaOverlap = NULL
AreaSp1 = NULL
AreaSp2 = NULL
Table = data.frame()

#For each species, split dataset 
spx <- split(Complete_Dataset$D13Ccor_consoI, list(Complete_Dataset$TaxaUsed),drop=FALSE)
spy <- split(Complete_Dataset$TP, list(Complete_Dataset$TaxaUsed),drop=FALSE)

# Run command: with crayfish
for (i in names(spx[c(1,3:5)])){
Overlap <- siar::overlap(spx$Crayfish, spy$Crayfish,
                         spx[[i]], spy[[i]], steps=1)
AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
Species1 <- append(Species1, "Crayfish", after = length(Species1))
Species2 <- append(Species2, names(spx[i]), after = length(Species2))
AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
} 
 
# Run command: with agile frog
for (i in names(spx[c(3:5)])){
Overlap <- siar::overlap(spx$`Agile frog`, spy$`Agile frog`,
                         spx[[i]], spy[[i]], steps=1)
AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
Species1 <- append(Species1, "Agile frog", after = length(Species1))
Species2 <- append(Species2, names(spx[i]), after = length(Species2))
AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
} 

# Run command: with marbled newt
for (i in names(spx[c(4:5)])){
Overlap <- siar::overlap(spx$`Marbled newt`, spy$`Marbled newt`,
                         spx[[i]], spy[[i]], steps=1)
AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
Species1 <- append(Species1, "Marbled newt", after = length(Species1))
Species2 <- append(Species2, names(spx[i]), after = length(Species2))
AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
}


# Run command: with palmate newt/tree frog
Overlap <- siar::overlap(spx$`Palmate newt`, spy$`Palmate newt`,
                         spx$`Tree frog`, spy$`Tree frog`, steps=1)
AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
Species1 <- append(Species1, "Palmate newt", after = length(Species1))
Species2 <- append(Species2, "Tree frog", after = length(Species2))
AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))

```


Analyses of overlap areas
``` {r analyses of overlaps}

Table<-cbind(Species1, Species2, AreaOverlap = round(AreaOverlap,4), 
             AreaSp1 = round(AreaSp1,4), AreaSp2 = round(AreaSp2,4))

Table <- as.data.frame(Table)
Table$Prct_overlap <- round(as.numeric(Table$AreaOverlap) / (as.numeric(Table$AreaSp1)+as.numeric(Table$AreaSp2)),4)
Table$Prct_overlap_sp2 <- round(as.numeric(Table$AreaOverlap) / as.numeric(Table$AreaSp2),4)
Table$Prct_overlap_sp1 <- round(as.numeric(Table$AreaOverlap) / as.numeric(Table$AreaSp1),4)
Table

write.csv(Table, file.path(here(), "exported_data", "Overlap_species_complete.csv"), row.names = F)
```






# 2- Population metrics

We calculate population metrics for amphibians and crayfish.

We begin with selecting the species of interest and removing populations with inadequate samples, i.e. large populations with n < 10. Some populations have n < 10 but are kept because we have good indicators from CMR that the sample size is representative of the population.

```{r select appropriate populations, include=FALSE}

# Pop_Dataset <- Dataset %>% dplyr::filter(!Group == "C1")
Pop_Dataset_Complete <- Dataset_complete %>% dplyr::filter(!Group == "C1")
# dim(Pop_Dataset)[1]
dim(Pop_Dataset_Complete)
# Keeping the outliers does not change the list of populations with n < 10


# We want to exclude target populations with n < 10, except three pops: marbled newt in ponds 56 and 94, and palmate newt in pop 112
# Pop_Dataset %>%  group_by(TaxaUsed, Pond) %>% summarise(n = n()) %>% filter(n < 10)
# 
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Crayfish"|!Pop_Dataset$Pond=="70",]
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Agile frog"|!Pop_Dataset$Pond=="22",]
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Agile frog"|!Pop_Dataset$Pond== "82_2016",]
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Marbled newt"|!Pop_Dataset$Pond== "111",]
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Marbled newt"|!Pop_Dataset$Pond== "112",]
# #Marbled 56 kept
# #Marbled 94 kept
# #Palmate 112 kept
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Palmate newt"|!Pop_Dataset$Pond=="72",]
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Palmate newt"|!Pop_Dataset$Pond== "82_2016",]
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Palmate newt"|!Pop_Dataset$Pond== "88",]
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Palmate newt"|!Pop_Dataset$Pond=="888",]
# Pop_Dataset<-Pop_Dataset[!Pop_Dataset$TaxaUsed=="Palmate newt"|!Pop_Dataset$Pond=="89",]
# 
# # The following do not apply since Crayfish juvenile and adult are merged together
# # Pop_Dataset<-Pop_Dataset[!Pop_Dataset$Taxa=="Juv crayfish"|!Pop_Dataset$Pond=="66",]
# # Pop_Dataset<-Pop_Dataset[!Pop_Dataset$Taxa=="Juv crayfish"|!Pop_Dataset$Pond=="70",]
# # Pop_Dataset<-Pop_Dataset[!Pop_Dataset$Taxa=="Juv crayfish"|!Pop_Dataset$Pond=="72",]
# # Pop_Dataset<-Pop_Dataset[!Pop_Dataset$Taxa=="Juv crayfish"|!Pop_Dataset$Pond== "82_2016",]
# # Pop_Dataset<-Pop_Dataset[!Pop_Dataset$Taxa=="Juv crayfish"|!Pop_Dataset$Pond=="888",]
# 
# table(Pop_Dataset$Pond, Pop_Dataset$TaxaUsed)


# Same thing with the complete dataset
Pop_Dataset_Complete %>%  group_by(TaxaUsed, Pond) %>% summarise(n = n()) %>% filter(n < 10)

Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Crayfish"|!Pop_Dataset_Complete$Pond=="70",]
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Agile frog"|!Pop_Dataset_Complete$Pond=="22",]
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Agile frog"|!Pop_Dataset_Complete$Pond=="82_2016",]
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Marbled newt"|!Pop_Dataset_Complete$Pond=="111",]
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Marbled newt"|!Pop_Dataset_Complete$Pond=="112",]
#Marbled 56 kept
#Marbled 94 kept
#Palmate 112 kept
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Palmate newt"|!Pop_Dataset_Complete$Pond=="72",]
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Palmate newt"|!Pop_Dataset_Complete$Pond== "82_2016",]
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Palmate newt"|!Pop_Dataset_Complete$Pond== "88",]
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Palmate newt"|!Pop_Dataset_Complete$Pond=="888",]
Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$TaxaUsed=="Palmate newt"|!Pop_Dataset_Complete$Pond=="89",]

# Juvenile crayfish do not apply
# Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$Taxa=="Juv crayfish"|!Pop_Dataset_Complete$Pond=="66",]
# Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$Taxa=="Juv crayfish"|!Pop_Dataset_Complete$Pond=="70",]
# Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$Taxa=="Juv crayfish"|!Pop_Dataset_Complete$Pond=="72",]
# Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$Taxa=="Juv crayfish"|!Pop_Dataset_Complete$Pond== "82_2016",]
# Pop_Dataset_Complete<-Pop_Dataset_Complete[!Pop_Dataset_Complete$Taxa=="Juv crayfish"|!Pop_Dataset_Complete$Pond=="888",]

table(Pop_Dataset_Complete$Pond, Pop_Dataset_Complete$TaxaUsed)

```

Check if the number of populations is correct

```{r check number of individuals is correct, include=FALSE}
# dim(Pop_Dataset)
# 
# dim(Pop_Dataset[Pop_Dataset$Taxa=="Agile frog",])[1] == 274 
# dim(Pop_Dataset[Pop_Dataset$Taxa=="Tree frog",])[1] == 186
# dim(Pop_Dataset[Pop_Dataset$Taxa=="Palmate newt",])[1] == 209
# dim(Pop_Dataset[Pop_Dataset$Taxa=="Marbled newt",])[1] == 101
# dim(Pop_Dataset[Pop_Dataset$TaxaUsed=="Crayfish",])[1] == 421 #88, J and K included

dim(Pop_Dataset_Complete[Pop_Dataset_Complete$Taxa=="Agile frog",])[1] == 274 
dim(Pop_Dataset_Complete[Pop_Dataset_Complete$Taxa=="Tree frog",])[1] == 187
dim(Pop_Dataset_Complete[Pop_Dataset_Complete$Taxa=="Palmate newt",])[1] == 209
dim(Pop_Dataset_Complete[Pop_Dataset_Complete$Taxa=="Marbled newt",])[1] == 101
dim(Pop_Dataset_Complete[Pop_Dataset_Complete$TaxaUsed=="Crayfish",])[1] == 422 #88, J and K incl

```

We first calculate the six metrics of Layman (TA, CR, NR, CD, MNND, SDNND) on each population for each pond. This is done by providing a vector containing the list of populations and ponds.


```{r calculate Layman metrics for populations, include=FALSE}
#Without outliers
Pop_Dataset

# UNCOMMENT THIS LINE AND RERUN THIS CHUNCK THE FOLLOWING LINE FOR THE METRICS WITH THE COMPLETE DATASET
Pop_Dataset <- Pop_Dataset_Complete

# 6 niches (2 for crayfish, 4 for amphibians)
spx <- split(Pop_Dataset$D13Ccor_consoI, 
             list(Pop_Dataset$Pond, Pop_Dataset$TaxaUsed), drop = TRUE)
length(spx)

spy <- split(Pop_Dataset$TP, list(Pop_Dataset$Pond, Pop_Dataset$TaxaUsed), drop = TRUE)
length(spy)

Table <- data.frame(metrics = c("dX_range", "dY_range", "NND", "SDNND", "CD", "TA"))

for (i in 1:length(spx)) {  
  Metrics <- laymanMetrics(spx[[i]],spy[[i]])
  Data <- as.data.frame(Metrics$metrics)
  Data$metrics <- rownames(Data)
  colnames(Data)[1] <- names(spx[i])
  Table <- merge(Table, Data, by = 'metrics', suffixes = colnames(Data)[1], all=TRUE)
}

Pop_metrics_Layman<-as.data.frame(t(Table))
colnames(Pop_metrics_Layman) <- Pop_metrics_Layman[1,]
Pop_metrics_Layman <- Pop_metrics_Layman[-1,]
Pop_metrics_Layman$Code <- rownames(Pop_metrics_Layman)


```


Now we calculate the equivalent using the bootstrapping method. For each pop, we sample 10 individuals in the population, calculate the 6 metrics of Layman, then do it again 999 times.

```{r calculate Layman metrics with bootstrapping for populations, include=FALSE}

# Make sure we find the same results everytime by setting the seed
set.seed(10)

# We group samples by pond and species again
spx <- split(Pop_Dataset$D13Ccor_consoI, 
             list(Pop_Dataset$Pond, Pop_Dataset$TaxaUsed), drop = TRUE)
length(spx)

spy <- split(Pop_Dataset$TP, list(Pop_Dataset$Pond, Pop_Dataset$TaxaUsed), drop = TRUE)
length(spy)

# Number of bootstrapped required
nr <- 9999 # the number of reps
ncol = 57

# Create empty vectors to store the results 
dNR <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
dCR <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
TA <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
CD <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
MNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
SDNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))

names(dNR)<-names(spx)
names(dCR)<-names(spx)
names(TA)<-names(spx)
names(CD)<-names(spx)
names(MNND)<-names(spx)
names(SDNND)<-names(spx)

for (j in 1:length(spx)) {
  if (length(spx[[j]]) > 10) {
    for (i in 1:nr) {
      simX<-sample(spx[[j]], 10, replace = FALSE, prob = NULL)
      simY<-sample(spy[[j]],10,replace = FALSE,prob=NULL)
      layman <- laymanmetrics(simX, simY)
      dNR[i,j] <- layman$dN_range
      dCR[i,j] <- layman$dC_range
      TA[i,j] <- layman$hull$TA
      CD[i,j] <- layman$CD
      MNND[i,j] <- layman$MNND
      SDNND[i,j] <- layman$SDNND
    }
    }
  }


```


```{r calculate average bootstrapped Layman metrics for populations, include=FALSE}

Pop_metrics_bootstrapped <- as.data.frame(cbind(NRb = colMeans(dNR), 
                                  CRb = colMeans(dCR), 
                                  TAb = colMeans(TA),
                                  CDb = colMeans(CD), 
                                  MNNDb = colMeans(MNND), 
                                  SDNNDb = colMeans(SDNND),
                                  Code = colnames(dNR)))

```


Finally, we calculate TA, SEA and SEAc for populations from Layman. 

```{r calculate population ellipses, include=FALSE}

Pop_Dataset_ellipses <- Pop_Dataset %>% 
  dplyr::select("iso1" = D13Ccor_consoI, "iso2" = TP, "group" = TaxaUsed, "community" = Pond) %>%
  as.data.frame()

Pop_dataset_SIBER <- createSiberObject(Pop_Dataset_ellipses)
Pop_metrics_SEAc <- groupMetricsML(Pop_dataset_SIBER)

#Table transformation
Pop_metrics_ellipses <- as.data.frame(t(Pop_metrics_SEAc))
Pop_metrics_ellipses$Code <- rownames(Pop_metrics_ellipses)
Pop_metrics_ellipses$Pond <- gsub("[.].*","",Pop_metrics_ellipses$Code)
Pop_metrics_ellipses$TaxaUsed <- gsub(".*.[.]","",Pop_metrics_ellipses$Code)

```


Finally, we merge the three datasets.


```{r merge datasets, include=FALSE}

Pop_metrics <- merge(Pop_metrics_Layman, Pop_metrics_bootstrapped, by = "Code")
dim(Pop_metrics)

#Create a composite metric to avoid NAs in bootstrapped metrics
Pop_metrics$NRcp = Pop_metrics$NRb 
Pop_metrics$CRcp = Pop_metrics$CRb
Pop_metrics$TAcp = Pop_metrics$TAb
Pop_metrics$CDcp = Pop_metrics$CDb
Pop_metrics$MNNDcp = Pop_metrics$MNNDb
Pop_metrics$SDNNDcp = Pop_metrics$SDNNDb

for (i in 1:length(Pop_metrics$NRb)) {
  if (is.na(Pop_metrics$NRb[i]) == TRUE) {
    Pop_metrics$NRcp[i] <- Pop_metrics$dY_range[i]
    Pop_metrics$CRcp[i] <- Pop_metrics$dX_range[i]
    Pop_metrics$TAcp[i] <- Pop_metrics$TA[i]
    Pop_metrics$CDcp[i] <- Pop_metrics$CD[i]
    Pop_metrics$MNNDcp[i] <- Pop_metrics$NND[i]
    Pop_metrics$SDNNDcp[i] <- Pop_metrics$SDNND[i]
  }
}

Pop_metrics <- merge(Pop_metrics, Pop_metrics_ellipses[-1], by = "Code")


ggplot(aes(x = as.numeric(CRcp), y = as.numeric(dX_range)), data = Pop_metrics) +
         geom_point() +
  geom_abline(intercept = 0, slope = 1)


ggplot(aes(x = as.numeric(NRcp), y = as.numeric(dY_range)), data = Pop_metrics) +
         geom_point() +
  geom_abline(intercept = 0, slope = 1)

# COMMENT THIS LINE AND UNCOMMENT THE NEXT ONE FOR THE COMPLETE DATASET
write.csv(Pop_metrics, file.path(here(), "exported_data", "Population_metrics_subset.csv"), row.names = F)
# write.csv(Pop_metrics, file.path(here(), "exported_data", "Population_metrics_complete.csv"), row.names = F)
```


```{r generate the long version of this table for Supp Mat}
Pop_metrics <- read.csv(file.path(here(), "exported_data", "Population_metrics_complete.csv"))
Mean_position <- read.csv(file.path(here(), "exported_data", "Meanposition_SEAc_complete.csv"))
names(Mean_position)[3] = "Code"
Mean_position %<>% select(-Pond, -Species) 
Pop_metrics <- merge(Pop_metrics, Mean_position, by = "Code")
Pop_metrics %<>% select(Pond, TaxaUsed, CRcp, NRcp, TAcp, CDcp, MNNDcp, SDNNDcp, SEAc, TPavg, D13Cavg)

Pop_metrics_long <- rbind(Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "CRcp") %>% add_column(Metrics = "CRcp"),
      Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "NRcp") %>% add_column(Metrics = "NRcp"),
      Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "TAcp") %>% add_column(Metrics = "TAcp"),
      Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "CDcp") %>% add_column(Metrics = "CDcp"),
      Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "MNNDcp") %>% add_column(Metrics = "MNNDcp"),
      Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "SDNNDcp") %>% add_column(Metrics = "SDNNDcp"),
      Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "SEAc") %>% add_column(Metrics = "SEAc"),
      Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "TPavg") %>% add_column(Metrics = "TPavg"),
      Pop_metrics %>% dcast(Pond ~ TaxaUsed, value.var = "D13Cavg") %>% add_column(Metrics = "D13Cavg"))
      

write.csv(Pop_metrics_long, file.path(here(), "exported_data", "Pop_metrics_complete_long.csv"), row.names = F)
```


## Compare complete and w/o outliers
```{r comparison}
Popmetrics <- read.csv(file.path(here(), "exported_data", "Population_metrics_subset.csv"), header=T)
Popmetrics_complete <- read.csv(file.path(here(), "exported_data", "Population_metrics_complete.csv"), header=T)

cbind(Popmetrics$Code,
      round(Popmetrics$NRcp - Popmetrics_complete$NRcp, 1),
      round(Popmetrics$CRcp - Popmetrics_complete$CRcp, 1),
      round(Popmetrics$TAcp - Popmetrics_complete$TAcp,1),
      round(Popmetrics$CDcp - Popmetrics_complete$CDcp,1),
      round(Popmetrics$MNNDcp - Popmetrics_complete$MNNDcp,1),
      round(Popmetrics$SDNNDcp - Popmetrics_complete$SDNNDcp,1),
      round(Popmetrics$SEA - Popmetrics_complete$SEA,1),
      round(Popmetrics$SEAc - Popmetrics_complete$SEAc,1))

# The only difference is for RV22 and ECR82 (TPAL82 has been removed anyway because n<10)
# MONITOR THESE POPULATIONS IN FURTHER ANALYZES

```


## Calculate CD between populations and global species niche
Calculate Layman metrics - Agile frog pop vs global
```{r calculate Layman metrics for agile frog, include=FALSE}

# Select agile frog
head(Pop_Dataset_Complete)
Agile_dataset <- Pop_Dataset_Complete %>% filter(TaxaUsed == "Agile frog")
head(Agile_dataset)

# Create ponds
spx <- split(Agile_dataset$D13Ccor_consoI, Agile_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Agile_dataset$TP, Agile_dataset$Pond, drop = TRUE)
length(spy)

Table <- data.frame(metrics = c("dX_range", "dY_range", "NND", "SDNND", "CD", "TA"))

for (i in 1:length(spx)) {  
  Metrics <- laymanMetrics(spx[[i]],spy[[i]])
  Data <- as.data.frame(Metrics$metrics)
  Data$metrics <- rownames(Data)
  colnames(Data)[1] <- names(spx[i])
  Table <- merge(Table, Data, by = 'metrics', suffixes = colnames(Data)[1], all=TRUE)
}

Layman_agile <- as.data.frame(t(Table))
colnames(Layman_agile) <- Layman_agile[1,]
Layman_agile <- Layman_agile[-1,]
Layman_agile$Code <- rownames(Layman_agile)
Layman_agile %<>% mutate(Species = "Agile frog") 

```

Calculate Layman metrics - Tree frog pop vs global
```{r calculate Layman metrics for tree frog, include=FALSE}

# Select agile frog
head(Pop_Dataset_Complete)
Treefrog_dataset <- Pop_Dataset_Complete %>% filter(TaxaUsed == "Tree frog")
head(Treefrog_dataset)

# Create ponds
spx <- split(Treefrog_dataset$D13Ccor_consoI, Treefrog_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Treefrog_dataset$TP, Treefrog_dataset$Pond, drop = TRUE)
length(spy)

Table <- data.frame(metrics = c("dX_range", "dY_range", "NND", "SDNND", "CD", "TA"))

for (i in 1:length(spx)) {  
  Metrics <- laymanMetrics(spx[[i]],spy[[i]])
  Data <- as.data.frame(Metrics$metrics)
  Data$metrics <- rownames(Data)
  colnames(Data)[1] <- names(spx[i])
  Table <- merge(Table, Data, by = 'metrics', suffixes = colnames(Data)[1], all=TRUE)
}

Layman_treefrog <- as.data.frame(t(Table))
colnames(Layman_treefrog) <- Layman_treefrog[1,]
Layman_treefrog <- Layman_treefrog[-1,]
Layman_treefrog$Code <- rownames(Layman_treefrog)
Layman_treefrog %<>% mutate(Species = "Tree frog") 
```

Calculate Layman metrics - Palmate newt pop vs global
```{r calculate Layman metrics for palmate newt, include=FALSE}

# Select agile frog
head(Pop_Dataset_Complete)
Palmate_dataset <- Pop_Dataset_Complete %>% filter(TaxaUsed == "Palmate newt")
head(Palmate_dataset)

# Create ponds
spx <- split(Palmate_dataset$D13Ccor_consoI, Palmate_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Palmate_dataset$TP, Palmate_dataset$Pond, drop = TRUE)
length(spy)

Table <- data.frame(metrics = c("dX_range", "dY_range", "NND", "SDNND", "CD", "TA"))

for (i in 1:length(spx)) {  
  Metrics <- laymanMetrics(spx[[i]],spy[[i]])
  Data <- as.data.frame(Metrics$metrics)
  Data$metrics <- rownames(Data)
  colnames(Data)[1] <- names(spx[i])
  Table <- merge(Table, Data, by = 'metrics', suffixes = colnames(Data)[1], all=TRUE)
}

Layman_palmate <- as.data.frame(t(Table))
colnames(Layman_palmate) <- Layman_palmate[1,]
Layman_palmate <- Layman_palmate[-1,]
Layman_palmate$Code <- rownames(Layman_palmate)
Layman_palmate %<>% mutate(Species = "Palmate newt") 
```

Calculate Layman metrics - Palmate newt pop vs global
```{r calculate Layman metrics for palmate newt, include=FALSE}

# Select agile frog
head(Pop_Dataset_Complete)
Marbled_dataset <- Pop_Dataset_Complete %>% filter(TaxaUsed == "Marbled newt")
head(Marbled_dataset)

# Create ponds
spx <- split(Marbled_dataset$D13Ccor_consoI, Marbled_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Marbled_dataset$TP, Marbled_dataset$Pond, drop = TRUE)
length(spy)

Table <- data.frame(metrics = c("dX_range", "dY_range", "NND", "SDNND", "CD", "TA"))

for (i in 1:length(spx)) {  
  Metrics <- laymanMetrics(spx[[i]],spy[[i]])
  Data <- as.data.frame(Metrics$metrics)
  Data$metrics <- rownames(Data)
  colnames(Data)[1] <- names(spx[i])
  Table <- merge(Table, Data, by = 'metrics', suffixes = colnames(Data)[1], all=TRUE)
}

Layman_marbled <- as.data.frame(t(Table))
colnames(Layman_marbled) <- Layman_marbled[1,]
Layman_marbled <- Layman_marbled[-1,]
Layman_marbled$Code <- rownames(Layman_marbled)
Layman_marbled %<>% mutate(Species = "Marbled newt") 
```

Calculate Layman metrics - crayfish pop vs global
```{r calculate Layman metrics for crayfish, include=FALSE}

# Select agile frog
head(Pop_Dataset_Complete)
Crayfish_dataset <- Pop_Dataset_Complete %>% filter(TaxaUsed == "Crayfish")
head(Crayfish_dataset)

# Create ponds
spx <- split(Crayfish_dataset$D13Ccor_consoI, Crayfish_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Crayfish_dataset$TP, Crayfish_dataset$Pond, drop = TRUE)
length(spy)

Table <- data.frame(metrics = c("dX_range", "dY_range", "NND", "SDNND", "CD", "TA"))

for (i in 1:length(spx)) {  
  Metrics <- laymanMetrics(spx[[i]],spy[[i]])
  Data <- as.data.frame(Metrics$metrics)
  Data$metrics <- rownames(Data)
  colnames(Data)[1] <- names(spx[i])
  Table <- merge(Table, Data, by = 'metrics', suffixes = colnames(Data)[1], all=TRUE)
}

Layman_crayfish <- as.data.frame(t(Table))
colnames(Layman_crayfish) <- Layman_crayfish[1,]
Layman_crayfish <- Layman_crayfish[-1,]
Layman_crayfish$Code <- rownames(Layman_crayfish)
Layman_crayfish %<>% mutate(Species = "Crayfish") 

```


## Bootstrapped version
Now we calculate the equivalent using the bootstrapping method. For each pop, we sample 10 individuals in the population, calculate the 6 metrics of Layman, then do it again 999 times.

Agile frog
```{r calculate Layman metrics with bootstrapping for agile frog, include=FALSE}

# Make sure we find the same results everytime by setting the seed
set.seed(10)

# We group samples by pond and species again
spx <- split(Agile_dataset$D13Ccor_consoI, 
             Agile_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Agile_dataset$TP,
             Agile_dataset$Pond, drop = TRUE)
length(spy)

# Number of bootstrapped required
nr <- 9999 # the number of reps
ncol = length(spy)

# Create empty vectors to store the results 
CD <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
MNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
SDNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))

names(CD)<-names(spx)
names(MNND)<-names(spx)
names(SDNND)<-names(spx)

for (j in 1:length(spx)) {
  if (length(spx[[j]]) > 10) {
    for (i in 1:nr) {
      simX<-sample(spx[[j]], 10, replace = FALSE, prob = NULL)
      simY<-sample(spy[[j]],10,replace = FALSE,prob=NULL)
      layman <- laymanmetrics(simX, simY)
      CD[i,j] <- layman$CD
      MNND[i,j] <- layman$MNND
      SDNND[i,j] <- layman$SDNND
    }
    }
}

Agile_bootstrapped <- as.data.frame(cbind(
                                  CDb = colMeans(CD), 
                                  MNNDb = colMeans(MNND), 
                                  SDNNDb = colMeans(SDNND),
                                  Code = colnames(CD)))

```

Tree frog
```{r calculate Layman metrics with bootstrapping for tree frog, include=FALSE}

# Make sure we find the same results everytime by setting the seed
set.seed(10)

# We group samples by pond and species again
spx <- split(Treefrog_dataset$D13Ccor_consoI, 
             Treefrog_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Treefrog_dataset$TP,
             Treefrog_dataset$Pond, drop = TRUE)
length(spy)

# Number of bootstrapped required
nr <- 9999 # the number of reps
ncol = length(spy)

# Create empty vectors to store the results 
CD <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
MNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
SDNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))

names(CD)<-names(spx)
names(MNND)<-names(spx)
names(SDNND)<-names(spx)

for (j in 1:length(spx)) {
  if (length(spx[[j]]) > 10) {
    for (i in 1:nr) {
      simX<-sample(spx[[j]], 10, replace = FALSE, prob = NULL)
      simY<-sample(spy[[j]],10,replace = FALSE,prob=NULL)
      layman <- laymanmetrics(simX, simY)
      CD[i,j] <- layman$CD
      MNND[i,j] <- layman$MNND
      SDNND[i,j] <- layman$SDNND
    }
    }
}

Treefrog_bootstrapped <- as.data.frame(cbind(
                                  CDb = colMeans(CD), 
                                  MNNDb = colMeans(MNND), 
                                  SDNNDb = colMeans(SDNND),
                                  Code = colnames(CD)))

```

Palmate newt
```{r calculate Layman metrics with bootstrapping for palmate newt, include=FALSE}

# Make sure we find the same results everytime by setting the seed
set.seed(10)

# We group samples by pond and species again
spx <- split(Palmate_dataset$D13Ccor_consoI, 
             Palmate_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Palmate_dataset$TP,
             Palmate_dataset$Pond, drop = TRUE)
length(spy)

# Number of bootstrapped required
nr <- 9999 # the number of reps
ncol = length(spy)

# Create empty vectors to store the results 
CD <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
MNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
SDNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))

names(CD)<-names(spx)
names(MNND)<-names(spx)
names(SDNND)<-names(spx)

for (j in 1:length(spx)) {
  if (length(spx[[j]]) > 10) {
    for (i in 1:nr) {
      simX<-sample(spx[[j]], 10, replace = FALSE, prob = NULL)
      simY<-sample(spy[[j]],10,replace = FALSE,prob=NULL)
      layman <- laymanmetrics(simX, simY)
      CD[i,j] <- layman$CD
      MNND[i,j] <- layman$MNND
      SDNND[i,j] <- layman$SDNND
    }
    }
}

Palmate_bootstrapped <- as.data.frame(cbind(
                                  CDb = colMeans(CD), 
                                  MNNDb = colMeans(MNND), 
                                  SDNNDb = colMeans(SDNND),
                                  Code = colnames(CD)))

```

Marbled newt
```{r calculate Layman metrics with bootstrapping for marbled newt, include=FALSE}

# Make sure we find the same results everytime by setting the seed
set.seed(10)

# We group samples by pond and species again
spx <- split(Marbled_dataset$D13Ccor_consoI, 
             Marbled_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Marbled_dataset$TP,
             Marbled_dataset$Pond, drop = TRUE)
length(spy)

# Number of bootstrapped required
nr <- 9999 # the number of reps
ncol = length(spy)

# Create empty vectors to store the results 
CD <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
MNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
SDNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))

names(CD)<-names(spx)
names(MNND)<-names(spx)
names(SDNND)<-names(spx)

for (j in 1:length(spx)) {
  if (length(spx[[j]]) > 10) {
    for (i in 1:nr) {
      simX<-sample(spx[[j]], 10, replace = FALSE, prob = NULL)
      simY<-sample(spy[[j]],10,replace = FALSE,prob=NULL)
      layman <- laymanmetrics(simX, simY)
      CD[i,j] <- layman$CD
      MNND[i,j] <- layman$MNND
      SDNND[i,j] <- layman$SDNND
    }
    }
}

Marbled_bootstrapped <- as.data.frame(cbind(
                                  CDb = colMeans(CD), 
                                  MNNDb = colMeans(MNND), 
                                  SDNNDb = colMeans(SDNND),
                                  Code = colnames(CD)))

```

Crayfish
```{r calculate Layman metrics with bootstrapping for crayfish, include=FALSE}

# Make sure we find the same results everytime by setting the seed
set.seed(10)

# We group samples by pond and species again
spx <- split(Crayfish_dataset$D13Ccor_consoI, 
             Crayfish_dataset$Pond, drop = TRUE)
length(spx)

spy <- split(Crayfish_dataset$TP,
             Crayfish_dataset$Pond, drop = TRUE)
length(spy)

# Number of bootstrapped required
nr <- 9999 # the number of reps
ncol = length(spy)

# Create empty vectors to store the results 
CD <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
MNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))
SDNND <- data.frame(matrix(NA, ncol = ncol, nrow = nr))

names(CD)<-names(spx)
names(MNND)<-names(spx)
names(SDNND)<-names(spx)

for (j in 1:length(spx)) {
  if (length(spx[[j]]) > 10) {
    for (i in 1:nr) {
      simX<-sample(spx[[j]], 10, replace = FALSE, prob = NULL)
      simY<-sample(spy[[j]],10,replace = FALSE,prob=NULL)
      layman <- laymanmetrics(simX, simY)
      CD[i,j] <- layman$CD
      MNND[i,j] <- layman$MNND
      SDNND[i,j] <- layman$SDNND
    }
    }
}

Crayfish_bootstrapped <- as.data.frame(cbind(
                                  CDb = colMeans(CD), 
                                  MNNDb = colMeans(MNND), 
                                  SDNNDb = colMeans(SDNND),
                                  Code = colnames(CD)))

```


## Merge all datasets
```{r merge CD tables to create CDcp}

CD_global <- rbind(
  merge(Layman_agile %>% dplyr::select(Code, CD, NND, SDNND, Species), Agile_bootstrapped, by = "Code"),
  merge(Layman_treefrog %>% dplyr::select(Code, CD, NND, SDNND, Species), Treefrog_bootstrapped, by = "Code"),
  merge(Layman_palmate %>% dplyr::select(Code, CD, NND, SDNND, Species), Palmate_bootstrapped, by = "Code"),
  merge(Layman_marbled %>% dplyr::select(Code, CD, NND, SDNND, Species), Marbled_bootstrapped, by = "Code"),
  merge(Layman_crayfish %>% dplyr::select(Code, CD, NND, SDNND, Species), Crayfish_bootstrapped, by = "Code")
)

#Create a composite metric to avoid NAs in bootstrapped metrics
CD_global %<>% mutate(CDcp = CDb,
                      MNNDcp = MNNDb,
                      SDNNDcp = SDNNDb)


for (i in 1:length(CD_global$CDb)) {
  if (is.na(CD_global$CDb[i]) == TRUE) {
    CD_global$CDcp[i] <- CD_global$CD[i]
    CD_global$MNNDcp[i] <- CD_global$NND[i]
    CD_global$SDNNDcp[i] <- CD_global$SDNND[i]
  }
}

CD_global

write.csv(CD_global, file.path(here(), "exported_data", "CD_global.csv"), row.names = F)
```



# 3- Coordinates of the standard ellipses (SEAc) 

Remember to remove n < 10 first (see part 2)

```{r merge datasets, include=FALSE}

# First
Pop_Dataset

# UNCOMMENT THIS LINE FOR THE COMPLETE DATASET
# Pop_Dataset = Pop_Dataset_Complete


table(Pop_Dataset$TaxaUsed)
spx <- split(Pop_Dataset$D13Ccor_consoI, list(Pop_Dataset$Pond, Pop_Dataset$TaxaUsed),drop=TRUE)
spy <- split(Pop_Dataset$TP, list(Pop_Dataset$Pond, Pop_Dataset$TaxaUsed),drop=TRUE)
length(spx)

Pop<-NULL
coordY<-NULL
coordX<-NULL
j=1

for (i in 1:length(spx)) {
  SE <- standard.ellipse(spx[[i]],spy[[i]],steps=1)
  Pop[j:(j+360)]<-names(spx)[i]
  coordY[j:(j+360)]<-SE$ySEAc[1:361]
  coordX[j:(j+360)]<-SE$xSEAc[1:361]
  j=j+361
}

CoordSEAc<-cbind(Pop,coordY,coordX)
summary(CoordSEAc)
CoordSEAc<-as.data.frame(CoordSEAc)
CoordSEAc$coordY<-as.numeric(as.character(CoordSEAc$coordY))
CoordSEAc$coordX<-as.numeric(as.character(CoordSEAc$coordX))
str(CoordSEAc)
CoordSEAc$Pond<-gsub("\\..*","",CoordSEAc$Pop)
CoordSEAc$Species<-gsub(".*\\.","",CoordSEAc$Pop)
head(CoordSEAc)
CoordSEAc$Pond<-as.factor(CoordSEAc$Pond)
levels(CoordSEAc$Pond)

# COMMENT THIS LINE AND UNCOMMENT THE NEXT ONE FOR THE COMPLETE DATASET
write.csv(CoordSEAc, file.path(here(), "exported_data", "Coordinates_SEAc_subset.csv"), row.names = F)
# write.csv(CoordSEAc, file.path(here(), "exported_data", "Coordinates_SEAc_complete.csv"), row.names = F)
``` 


```{r centroides of standard ellipses}

CoordSEAc <- read.csv(file.path(here(), "exported_data", "Coordinates_SEAc_subset.csv"), header=T)
CoordSEAc_complete <- read.csv(file.path(here(), "exported_data", "Coordinates_SEAc_complete.csv"), header=T)


Mean_position <- CoordSEAc %>% group_by(Pond, Species, Pop) %>%
  summarise(TPavg = mean(coordY),
            D13Cavg = mean(coordX))

dim(Mean_position)[1]

write.csv(Mean_position, file.path(here(), "exported_data", "Meanposition_SEAc.csv"), row.names = F)



Mean_position <- CoordSEAc_complete %>% group_by(Pond, Species, Pop) %>%
  summarise(TPavg = mean(coordY),
            D13Cavg = mean(coordX))

dim(Mean_position)[1]

write.csv(Mean_position, file.path(here(), "exported_data", "Meanposition_SEAc_complete.csv"), row.names = F)
```




#4- Area of ellipses overlap

````{r area of ellipses overlap: initialize vectors}

Pop_Dataset

# UNCOMMENT THIS LINE FOR THE COMPLETE DATASET
# Pop_Dataset <- Pop_Dataset_Complete

# WARNING: Crayfish juv and ad don't work because populations with n < 10 unless removed earlier 

#Split dataset by species
# AdCrayfish <- Pop_Dataset[Pop_Dataset$Taxa=="Ad crayfish",]
# JuvCrayfish <- Pop_Dataset[Pop_Dataset$Taxa=="Juv crayfish",]
Crayfish <- Pop_Dataset %>% filter(TaxaUsed == "Crayfish")
Treefrog <- Pop_Dataset[Pop_Dataset$Taxa=="Tree frog",] 
Agilefrog <- Pop_Dataset[Pop_Dataset$Taxa=="Agile frog",] 
Palmatenewt <- Pop_Dataset[Pop_Dataset$Taxa=="Palmate newt",] 
Marblednewt <- Pop_Dataset[Pop_Dataset$Taxa=="Marbled newt",] 


#For each species, split dataset by pond
# spx_AdCrayfish <- split(AdCrayfish$D13Ccor_consoI, list(AdCrayfish$Pond),drop=FALSE)
# spy_AdCrayfish <- split(AdCrayfish$TP, list(AdCrayfish$Pond),drop=FALSE)
# spx_JuvCrayfish <- split(JuvCrayfish$D13Ccor_consoI, list(JuvCrayfish$Pond),drop=FALSE)
# spy_JuvCrayfish <- split(JuvCrayfish$TP, list(JuvCrayfish$Pond),drop=FALSE)
spx_Agilefrog <- split(Agilefrog$D13Ccor_consoI, list(Agilefrog$Pond),drop=FALSE)
spy_Agilefrog <- split(Agilefrog$TP, list(Agilefrog$Pond),drop=FALSE)
spx_Treefrog <- split(Treefrog$D13Ccor_consoI, list(Treefrog$Pond),drop=FALSE)
spy_Treefrog <- split(Treefrog$TP, list(Treefrog$Pond),drop=FALSE)
spx_Palmatenewt <- split(Palmatenewt$D13Ccor_consoI, list(Palmatenewt$Pond),drop=FALSE)
spy_Palmatenewt <- split(Palmatenewt$TP, list(Palmatenewt$Pond),drop=FALSE)
spx_Marblednewt <- split(Marblednewt$D13Ccor_consoI, list(Marblednewt$Pond),drop=FALSE)
spy_Marblednewt <- split(Marblednewt$TP, list(Marblednewt$Pond),drop=FALSE)

spx_Crayfish <- split(Crayfish$D13Ccor_consoI, list(Crayfish$Pond),drop=FALSE)
spy_Crayfish <- split(Crayfish$TP, list(Crayfish$Pond),drop=FALSE)

#Initialize vectors
Pond = NULL
Species1 = NULL
Species2 = NULL
AreaOverlap = NULL
AreaSp1 = NULL
AreaSp2 = NULL
Table = data.frame()
```


## Crayfish adult (DEPRECATED)
```{r area of ellipses overlap: loop over crayfish adults}

# for (i in names(spx_AdCrayfish)) {
#   if (length(spx_Agilefrog[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Ad crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Agile frog", after = length(Species2))
#     Overlap <- siar::overlap(spx_AdCrayfish[[i]], spy_AdCrayfish[[i]],
#                              spx_Agilefrog[[i]], spy_Agilefrog[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
#   
#   if (length(spx_Treefrog[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Ad crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Tree frog", after = length(Species2))
#     Overlap <- siar::overlap(spx_AdCrayfish[[i]], spy_AdCrayfish[[i]],
#                              spx_Treefrog[[i]], spy_Treefrog[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
#   
#   if (length(spx_Palmatenewt[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Ad crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Palmate newt", after = length(Species2))
#     Overlap <- siar::overlap(spx_AdCrayfish[[i]], spy_AdCrayfish[[i]],
#                              spx_Palmatenewt[[i]], spy_Palmatenewt[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
#   
#   if (length(spx_Marblednewt[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Ad crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Marbled newt", after = length(Species2))
#     Overlap <- siar::overlap(spx_AdCrayfish[[i]], spy_AdCrayfish[[i]],
#                              spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
#   
#   
#   if (length(spx_JuvCrayfish[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Ad crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Juv crayfish", after = length(Species2))
#     Overlap <- siar::overlap(spx_AdCrayfish[[i]], spy_AdCrayfish[[i]],
#                              spx_JuvCrayfish[[i]], spy_JuvCrayfish[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
# }



```


## Crayfish juveniles (DEPRECATED)
```{r area of ellipses overlap: loop over crayfish juveniles}

# for (i in names(spx_JuvCrayfish)) {
#   if (length(spx_Agilefrog[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Ad crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Agile frog", after = length(Species2))
#     Overlap <- siar::overlap(spx_JuvCrayfish[[i]], spy_JuvCrayfish[[i]],
#                              spx_Agilefrog[[i]], spy_Agilefrog[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
#   
#   if (length(spx_Treefrog[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Juv crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Tree frog", after = length(Species2))
#     Overlap <- siar::overlap(spx_JuvCrayfish[[i]], spy_JuvCrayfish[[i]],
#                              spx_Treefrog[[i]], spy_Treefrog[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
#   
#   if (length(spx_Palmatenewt[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Juv crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Palmate newt", after = length(Species2))
#     Overlap <- siar::overlap(spx_JuvCrayfish[[i]], spy_JuvCrayfish[[i]],
#                              spx_Palmatenewt[[i]], spy_Palmatenewt[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
#   
#   if (length(spx_Marblednewt[[i]]) != 0) {
#     Pond <- append(Pond, i, after = length(Pond))
#     Species1 <- append(Species1, "Juv crayfish", after = length(Species1))
#     Species2 <- append(Species2, "Marbled newt", after = length(Species2))
#     Overlap <- siar::overlap(spx_JuvCrayfish[[i]], spy_JuvCrayfish[[i]],
#                              spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
#     AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
#     AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
#     AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
#   }
# }

```

## All crayfish
```{r loop over all crayfish}

for (i in names(spx_Crayfish)) {
  if (length(spx_Agilefrog[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Crayfish", after = length(Species1))
    Species2 <- append(Species2, "Agile frog", after = length(Species2))
    Overlap <- siar::overlap(spx_Crayfish[[i]], spy_Crayfish[[i]],
                             spx_Agilefrog[[i]], spy_Agilefrog[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Treefrog[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Crayfish", after = length(Species1))
    Species2 <- append(Species2, "Tree frog", after = length(Species2))
    Overlap <- siar::overlap(spx_Crayfish[[i]], spy_Crayfish[[i]],
                             spx_Treefrog[[i]], spy_Treefrog[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Palmatenewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Crayfish", after = length(Species1))
    Species2 <- append(Species2, "Palmate newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Crayfish[[i]], spy_Crayfish[[i]],
                             spx_Palmatenewt[[i]], spy_Palmatenewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Marblednewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Crayfish", after = length(Species1))
    Species2 <- append(Species2, "Marbled newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Crayfish[[i]], spy_Crayfish[[i]],
                             spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
}
```

## Between amphibians
```{r area of ellipses overlap: loop between amphibians}

for (i in names(spx_Agilefrog)) {
  if (length(spx_Treefrog[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Agile frog", after = length(Species1))
    Species2 <- append(Species2, "Tree frog", after = length(Species2))
    Overlap <- siar::overlap(spx_Agilefrog[[i]], spy_Agilefrog[[i]],
                             spx_Treefrog[[i]], spy_Treefrog[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Palmatenewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Agile frog", after = length(Species1))
    Species2 <- append(Species2, "Palmate newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Agilefrog[[i]], spy_Agilefrog[[i]],
                             spx_Palmatenewt[[i]], spy_Palmatenewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Marblednewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Agile frog", after = length(Species1))
    Species2 <- append(Species2, "Marbled newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Agilefrog[[i]], spy_Agilefrog[[i]],
                             spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
}



for (i in names(spx_Treefrog)) {
  if (length(spx_Palmatenewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Tree frog", after = length(Species1))
    Species2 <- append(Species2, "Palmate newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Treefrog[[i]], spy_Treefrog[[i]],
                             spx_Palmatenewt[[i]], spy_Palmatenewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
  
  if (length(spx_Marblednewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Tree frog", after = length(Species1))
    Species2 <- append(Species2, "Marbled newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Treefrog[[i]], spy_Treefrog[[i]],
                             spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
}


for (i in names(spx_Palmatenewt)) {
  if (length(spx_Marblednewt[[i]]) != 0) {
    Pond <- append(Pond, i, after = length(Pond))
    Species1 <- append(Species1, "Palmate newt", after = length(Species1))
    Species2 <- append(Species2, "Marbled newt", after = length(Species2))
    Overlap <- siar::overlap(spx_Palmatenewt[[i]], spy_Palmatenewt[[i]],
                             spx_Marblednewt[[i]], spy_Marblednewt[[i]], steps=1)
    AreaOverlap <- append(AreaOverlap, Overlap$overlap, after = length(AreaOverlap))
    AreaSp1 <- append(AreaSp1, Overlap$area1, after = length(AreaSp1))
    AreaSp2 <- append(AreaSp2, Overlap$area2, after = length(AreaSp2))
  }
}

```


## Analyses of overlap areas

``` {r analyses of overlaps}

Table<-cbind(Pond, Species1, Species2, AreaOverlap = round(AreaOverlap,4), 
             AreaSp1 = round(AreaSp1,4), AreaSp2 = round(AreaSp2,4))

Table <- as.data.frame(Table)
Table$Prct_overlap <- round(as.numeric(Table$AreaOverlap) / (as.numeric(Table$AreaSp1)+as.numeric(Table$AreaSp2)),4)
Table$Prct_overlap_sp2 <- round(as.numeric(Table$AreaOverlap) / as.numeric(Table$AreaSp2),4)
Table$Prct_overlap_sp1 <- round(as.numeric(Table$AreaOverlap) / as.numeric(Table$AreaSp1),4)
Table

# COMMENT THIS LINE AND UNCOMMENT THE NEXT ONE FOR THE COMPLETE DATASET
write.csv(Table, file.path(here(), "exported_data", "Overlap_subset.csv"), row.names = F)
# write.csv(Table, file.path(here(), "exported_data", "Overlap_complete.csv"), row.names = F)
```


