---
title: "#4: Network plots with ellipses"
author: 
- Nadege Belouard^[UMR CNRS EcoBio, UMR DECOD, nadege.belouard@gmail.com]
date: "23/03/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

This vignette draws all the plots related to the trophic networks. First, we load the different packages and datasets that we need.

```{r setup, warning = F, message = F}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(magrittr)
library(here)

# Points
Dataset <- read.csv(file.path(here(), "exported_data", 
                              "Isotope_data_standardized.csv"), header=T)

# Ellipses
#Per population
CoordSEAc_pop <- read.csv(file.path(here(), 
                                    "exported_data", 
                                    "Coordinates_SEAc_populations.csv"), 
                          header=T)
names(CoordSEAc_pop)[1] <- "Taxa"

#Per species
CoordSEAc_sp <- read.csv(file.path(here(), "exported_data", 
                                   "Coordinates_SEAc_regional.csv"), header=T)


# Average points
Avg_data_complete <- Dataset %>% group_by(Taxa, Pond) %>%
  summarise(TPavg = mean(TP),
            D13Cavg = mean(D13Ccor_consoI),
            TPsd = sd(TP),
            D13Csd = sd(D13Ccor_consoI))


```

We summarize each taxa by the average point and its SD to graphically represent the position and the variability in the isotopic signature of each taxa.


# Plots with ellipses per species

## Figure 1: ellipses per species with points
```{r plot network with ellipses per species and all individuals}

# With outliers
M_ind <- Dataset %>% filter (Group %in% c("Crayfish", "Amphibians"))
M_coord <- CoordSEAc_sp
M_ind$Species <- factor(M_ind$Species, levels = c("Crayfish",
                                                    "Agile frog",
                                                    "Tree frog",
                                                    "Palmate newt",
                                                    "Marbled newt"))

species_ellipses <- ggplot(data = M_ind, aes(x = D13Ccor_consoI, y = TP)) +
  geom_point(aes(col = Species), alpha = 0.7, shape = 16) +
  geom_path(data = M_coord, aes(coordX, coordY, col = Species), lwd = 1.5) +
  scale_color_manual(values = c("#E69F00", 
                                "#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab(bquote(''*delta^13~C["cor"]*'')) +
  guides(col = guide_legend("Species")) +
  ylab("Trophic position") +
  theme_classic() + 
  # theme(legend.justification = c(0, 1), legend.position = c(0, 1))
  theme(legend.position = "right")

species_ellipses

ggsave(species_ellipses, filename = file.path(here(), "figures", 
                                              "ellipses_species.jpg"), 
       height = 3.5, width = 5)
```


# Plots with ellipses per population per pond

## Figure 2: plots with ellipses per population with points
```{r plot ellipses per population with points}

M_ind$Species <- factor(M_ind$Species, levels = c("Crayfish",
                                                    "Agile frog",
                                                    "Tree frog",
                                                    "Palmate newt",
                                                    "Marbled newt"))

# Set limits for x values
blank_data <- data.frame(Pond = c("P01", "P01",
                                   "P02", "P02",
                                   "P03", "P03",
                                   "P04", "P04",
                                   "P05", "P05",
                                   "P06", "P06",
                                   "P07", "P07",
                                   "P08", "P08",
                                   "P09", "P09",
                                   "P10", "P10",
                                   "P11", "P11",
                                   "P12", "P12",
                                   "P13", "P13",
                                   "P14", "P14",
                                  "P15", "P15",
                                   "P16", "P16",
                                   "P17", "P17",
                                   "P18", "P18"
                                   ),
                         y = 1,
                         x = c(-1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -4, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -2, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5,
                               -1.5, 1.5
                               ))

ellipses_pop <- ggplot(data = M_ind, aes(D13Ccor_consoI, TP)) +
  theme_classic() +
  geom_point(aes(col = Species), alpha = 0.7) +
  geom_blank(data = blank_data, aes(x = x, y = y)) +
  geom_path(data = CoordSEAc_pop, aes(coordX, coordY, col = Species),
            size=1, linetype=1) +
  scale_color_manual(values = c("#E69F00", #orange 
                                "#0072B2", #darkblue 
                                "#009E73", #green 
                                "#56B4E9", # lightblue
                                "#CC79A7")) + #pink
  facet_wrap(~Pond, scales = "free_x", 
             ncol = 3) +
  xlab(bquote(''*delta^13~C["cor"]*'')) +
  ylab("Trophic position") +
  guides(col = guide_legend("Species")) + 
  theme(legend.position = "bottom")

ellipses_pop

ggsave(ellipses_pop, filename = file.path(here(), "figures",
                                          "ellipses_pop_points_portrait.jpg"),
       width = 10, height = 15)

```
