---
title: "Network plots with ellipses"
author: "Nadege Belouard"
date: "07/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This vignette draws all the plots related to the trophic networks. First, we load the different packages and datasets that we need.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(ggrepel)
library("rgeos")
library(dplyr)
library(magrittr)
library(here)

#### Points
# With all points
Dataset_homogen <- read.csv(file.path(here(), "exported_data", "Isotope_data_standardized_reduced.csv"), header=T)
dim(Dataset_homogen)

# Add the variable Taxa Used
for (i in 1:length(Dataset_homogen$Pond)) {
  if (Dataset_homogen$Taxa[i] %in% c("Agile frog", "Tree frog", "Palmate newt", "Marbled newt")) {
    Dataset_homogen$TaxaUsed[i] = Dataset_homogen$Taxa[i]
  } else if(Dataset_homogen$Taxa[i] %in% c("Ad crayfish", "Juv crayfish")) {
    Dataset_homogen$TaxaUsed[i] = "Crayfish"
  } else{
    Dataset_homogen$TaxaUsed[i] = "Other"
  }
}



# Without outliers
Dataset_minusoutliers <- read.csv(file.path(here(), "exported_data", "Isotope_data_minusoutliers.csv"), header=T)
dim(Dataset_minusoutliers)

# Add the variable TaxaUsed
for (i in 1:length(Dataset_minusoutliers$Pond)) {
  if (Dataset_minusoutliers$Taxa[i] %in% c("Agile frog", "Tree frog", "Palmate newt", "Marbled newt")) {
    Dataset_minusoutliers$TaxaUsed[i] = Dataset_minusoutliers$Taxa[i]
  } else if(Dataset_minusoutliers$Taxa[i] %in% c("Ad crayfish", "Juv crayfish")) {
    Dataset_minusoutliers$TaxaUsed[i] = "Crayfish"
  } else{
    Dataset_minusoutliers$TaxaUsed[i] = "Other"
  }
}



### Ellipses
#Per species
CoordSEAc_complete_species <- read.csv(file.path(here(), "exported_data", "Coordinates_SEAc_species_complete.csv"), header=T)
CoordSEAc_minusoutliers_species <- read.csv(file.path(here(), "exported_data", "Coordinates_SEAc_species_subset.csv"), header=T)
names(CoordSEAc_complete_species)[1] <- "TaxaUsed"
names(CoordSEAc_minusoutliers_species)[1] <- "TaxaUsed"

#Per population
CoordSEAc <- read.csv(file.path(here(), "exported_data", "Coordinates_SEAc_subset.csv"), header=T)
# names(CoordSEAc_consoI)[5] <- "TaxaUsed"
CoordSEAc_complete <- read.csv(file.path(here(), "exported_data", "Coordinates_SEAc_complete.csv"), header=T)



### Average points
Avg_data_complete <- Dataset_homogen %>% group_by(Taxa, Pond) %>%
  summarise(TPavg = mean(TP),
            D13Cavg = mean(D13Ccor_consoI),
            TPsd = sd(TP),
            D13Csd = sd(D13Ccor_consoI))

Avg_data_minusoutliers <- Dataset_minusoutliers %>% group_by(Taxa, Pond) %>%
  summarise(TPavg = mean(TP),
            D13Cavg = mean(D13Ccor_consoI),
            TPsd = sd(TP),
            D13Csd = sd(D13Ccor_consoI))

```

# Add colors (DEPRECATED)
```{r add colors}

# Legend colors
# Dataset homogen (complete)
# Species <- levels(droplevels(as.factor(Dataset_homogen$TaxaUsed)))
# ColorPalette<- data.frame(cbind(TaxaUsed = c("Agile frog", "Tree frog",
#                                          "Palmate newt", "Marbled newt",
#                                          "Crayfish", "Other"),
#                                 Color = c("#0072B2", "#009E73",
#                                           "#56B4E9", "#CC79A7",
#                                           "#E69F00", "black"),
#                                 ColorName = c("Blue", "Green",
#                                               "Light blue", "Pink",
#                                               "Orange", "Black")))
#   
# Dataset_homogen <- merge(Dataset_homogen, ColorPalette, by = "TaxaUsed")
# Dataset_minusoutliers <- merge(Dataset_minusoutliers, ColorPalette, by = "TaxaUsed")
# CoordSEAc_complete <- merge(CoordSEAc_complete, ColorPalette, by = "TaxaUsed")
# CoordSEAc_minusoutliers <- merge(CoordSEAc_minusoutliers, ColorPalette, by = "TaxaUsed")

```

We summarize each taxa by the average point and its SD to graphically represent the position and the variability in the isotopic signature of each taxa.


# Plots with ellipses per species

## Figure 1
```{r plot network with ellipses per species and all individuals, include=FALSE}

# With outliers
M_ind <- Dataset_homogen %>% filter (Group %in% c("Crayfish", "Amphibians"))
dim(M_ind)
M_coord <- CoordSEAc_complete_species
M_ind$TaxaUsed <- factor(M_ind$TaxaUsed, levels = c("Crayfish",
                                                    "Agile frog",
                                                    "Tree frog",
                                                    "Palmate newt",
                                                    "Marbled newt"))

species_ellipses <- ggplot(data = M_ind, aes(x = D13Ccor_consoI, y = TP)) +
  geom_point(aes(col = TaxaUsed), alpha = 0.7, shape = 16) +
  geom_path(data = M_coord, aes(coordX, coordY, col = TaxaUsed), lwd = 1.5) +
  scale_color_manual(values = c("#E69F00", 
                                "#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab(bquote(''*delta^13~C["cor"]*'')) +
  guides(col = guide_legend("Taxa")) +
  ylab("Trophic position") +
  theme_classic() + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))

ggsave(species_ellipses, filename = file.path(here(), "figures", "species_ellipsesv2.jpg"), height = 4, width = 6)


# figure with ellipses only
species_ellipses <- ggplot() +
  geom_path(data = M_coord, aes(coordX, coordY, col = TaxaUsed), lwd = 1.5) +
  scale_color_manual(values = c("#0072B2", "#E69F00",
                                          "#CC79A7", "#56B4E9", 
                                          "#009E73")) +
  xlab("D13Ccor") +
  ylab("Trophic position") +
  guides(col = guide_legend("Taxa")) +
  theme_classic()

ggsave(species_ellipses, filename = file.path(here(), "figures", "species_ellipses_only.jpg"), height = 4, width = 6)



# Without outliers
M_ind_out <- Dataset_minusoutliers %>% filter (Group %in% c("Crayfish", "Amphibians"))
dim(M_ind_out)
M_coord_out <- CoordSEAc_minusoutliers

species_ellipses <- ggplot(data = M_ind_out, aes(x = D13Ccor_consoI, y = TP)) +
  geom_point(aes(col = Color), alpha = 0.7, shape = 16) +
  geom_path(data = M_coord_out, aes(coordX, coordY, col = Color), lwd = 1.5) +
  scale_color_identity() +
  xlab("D13Ccor") +
  ylab("Trophic position") +
  theme_classic()


ggsave(species_ellipses, filename = file.path(here(), "figures", "species_ellipses_minus outliers.jpg"), height = 6, width = 6)




# Highlight difference between the two
species_ellipses <- ggplot(data = M_ind_out, aes(x = D13Ccor_consoI, y = TP)) +
  geom_point(alpha = 0.7, shape = 16, size = 3) +
  geom_point(data = M_ind, aes(col = Color), alpha = 0.7, shape = 16) +
  scale_color_identity() +
  xlab("D13Ccor") +
  ylab("Trophic position") +
  theme_classic()

dim(M_ind)
dim(M_ind_out)


```



# Plots with ellipses per population per pond
## Figure 2
```{r plot ellipses only, include=FALSE}

names(CoordSEAc_complete)
ponds_code <- read.table(file.path(here(), "pond_code.txt"), h=T)
CoordSEAc_complete <- left_join(CoordSEAc_complete, ponds_code)
CoordSEAc_complete$Pond_code <- factor(CoordSEAc_complete$Pond_code, levels = c("U1", "U2", "U3", "U4", "U5", "U6",
                                                                                "I1", "I2", "I3", "I4", "I5", "I6",
                                                                                "I7", "I8", "I9", "I10", "I11", "I12",
                                                                                "I13", "I14"))
M_ind$TaxaUsed <- factor(M_ind$TaxaUsed, levels = c("Crayfish",
                                                    "Agile frog",
                                                    "Tree frog",
                                                    "Palmate newt",
                                                    "Marbled newt"))
M_ind <- left_join(M_ind, ponds_code)
M_ind$Pond_code <- factor(M_ind$Pond_code, levels = c("U1", "U2", "U3", "U4", "U5", "U6",
                                                                                "I1", "I2", "I3", "I4", "I5", "I6",
                                                                                "I7", "I8", "I9", "I10", "I11", "I12",
                                                                                "I13", "I14"))

# Facet wrap ellipses
ellipses_pop <- ggplot() +
  theme_classic() +
  geom_path(data = CoordSEAc_complete, aes(coordX, coordY, col = Species), size=2, linetype=1) +
  scale_color_manual(values = c("#0072B2", "#E69F00",
                                          "#CC79A7", "#56B4E9", 
                                          "#009E73")) +
  facet_wrap(~Pond_code, scales = "free_x") +
  xlab(bquote(''*delta^13~C["cor"]*'')) +
  ylab("Trophic position")

ellipses_pop
ggsave(ellipses_pop, filename = file.path(here(), "figures", "ellipses_pop_complete.jpg"), width = 15, height = 10)


# Set limits for x values
blank_data <- data.frame(Pond_code = c("U1", "U1", 
                                   "U2", "U2", 
                                   "U3", "U3",
                                   "U4", "U4",
                                   "U5", "U5",
                                   "U6", "U6",
                                   "I1", "I1",
                                   "I2", "I2",
                                   "I3", "I3",
                                   "I4", "I4",
                                   "I5", "I5",
                                   "I6", "I6",
                                   "I7", "I7",
                                   "I8", "I8",
                                   "I9", "I9",
                                   "I10", "I10",
                                   "I11", "I11",
                                   "I12", "I12",
                                   "I13", "I13",
                                   "I14", "I14"),
                         y = 1,
                         x = c(-1.5, 1.5, 
                               -2, 1.5, 
                               -4, 1.5,
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5, 
                               -1.5, 1.5))
blank_data$Pond_code <- factor(blank_data$Pond_code, levels = c("U1", "U2", "U3", "U4", "U5", "U6",
                                                                                "I1", "I2", "I3", "I4", "I5", "I6",
                                                                                "I7", "I8", "I9", "I10", "I11", "I12",
                                                                                "I13", "I14"))

# Facet wrap ellipses and points
# Dataset_target <- Dataset_homogen %>% filter(!TaxaUsed == "Other")

ellipses_pop <- ggplot(data = M_ind, aes(D13Ccor_consoI, TP)) +
  theme_classic() +
  geom_point(aes(col = TaxaUsed), alpha = 0.7) +
  geom_path(data = CoordSEAc_complete, aes(coordX, coordY, col = Species), size=1, linetype=1) +
  
  scale_color_manual(values = c("#E69F00", 
                                "#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  geom_blank(data = blank_data, aes(x = x, y = y)) +
  facet_wrap(~Pond_code, scales = "free_x") +
  xlab(bquote(''*delta^13~C["cor"]*'')) +
  ylab("Trophic position") +
  guides(col = guide_legend("Taxa"))

ellipses_pop

ggsave(ellipses_pop, filename = file.path(here(), "figures", "ellipses_points_pop_completev4.jpg"), width = 15, height = 10)

```


# Plots with ellipses + points + C1 (DEPRECATED)
```{r plot standardized networks consoI, include=FALSE}
getwd()

for (i in levels(as.factor(Dataset_homogen$Pond))) {
  M_ind <- Dataset_homogen %>% filter(Pond == i) %>%
    filter (Group %in% c("Crayfish", "Amphibians"))
  M_avg <- Avg_data_std %>% filter(Pond == i) %>%
    filter (Taxa %in% c("Corbiculidae", "Physa", "Zooplankton", "Gammarus", "Corixidae","Asellus"))
    # filter(Taxa == "Other")
  M_coord <- CoordSEAc_consoI %>% filter(Pond == i)
  # M_coord_oneniche <- CoordSEAc_consoI_oneniche %>% filter(Pond == i)
  
f <- ggplot(data = M_avg, aes(x = D13Cavg, y = TPavg)) +
  theme_classic() +
  ggtitle(paste("Pond",i)) +
  geom_text_repel(data = M_avg,
  aes(x = D13Cavg, y = TPavg, label = Taxa),
  colour = "darkgrey", size = 3) +
  geom_errorbar(data = M_avg,
  aes(y = TPavg, ymin = TPavg - TPsd, ymax = TPavg + TPsd),
  colour = "darkgrey", width = 0.05) +
  geom_errorbarh(data = M_avg,
  aes(x = D13Cavg, xmin = D13Cavg - D13Csd, xmax = D13Cavg + D13Csd),
  colour = "darkgrey", height = 0.05) +
  geom_point(data = M_ind, aes(x = D13Ccor_consoI, y = TP, colour = Palette), size = 1) +
  geom_path(data = M_coord, aes(coordX, coordY, col = Palette), size=0.5, linetype=1) +
  # geom_path(data = M_coord_oneniche, aes(coordX, coordY, col = Palette), size=0.5, linetype=2) +
  scale_colour_identity() +
  xlab("D13Ccor") +
  ylab("Trophic position") +
  xlim(c(min(Dataset_homogen$D13Ccor_consoI), max(Dataset_homogen$D13Ccor_consoI))) +
  ylim(c(min(Dataset_homogen$TP), max(Dataset_homogen$TP)))


ggsave(f, filename = file.path(here(), "figures", "networks_ellipses", paste("M",i,"_ellipses.png",sep="")), width=5, height=4)
# ggsave(f, filename = file.path("../figures/networks_ellipses_oneniche", paste("M",i,"_ellipses.png",sep="")), width=5, height=4)
}

```
