---
title: "#6: Statistical models"
author: 
- Nadege Belouard^[UMR CNRS EcoBio, UMR DECOD, nadege.belouard@gmail.com]
date: "23/03/2023"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2
  html_document:
    toc: TRUE
    toc_depth: 3
params:
  show_code: FALSE
  export_figures: TRUE
editor_options: 
  chunk_output_type: console
---

This vignette tests the link between the isotopic niche metrics of amphibians and candidate variables, including crayfish.


# 1. Setup

We begin by loading the required libraries and datasets.

```{r setup, include = FALSE}

#Load packages
library(lme4)
library(car)
library(MuMIn)
library(ggplot2)
library(RVAideMemoire)
library(ggrepel)
library(magrittr)
library(Hmisc)
library(usdm)
library(tidyr)
library(dplyr)
library(tibble)
library(gridExtra)
library(here)

```

```{r load datasets}
#Load datasets
Envt <- read.csv(here::here("data_raw", "Raw_environment.csv"), h = T)
Envt %<>% mutate(Crayfish01 = ifelse(CrayfishCPUE == 0, 0, 1))
         
# Dataset
Niche_metrics <- read.csv(file.path(here(), "exported_data", 
                                    "Population_metrics.csv"), h = T) %>% 
  dplyr::select(Code, Pond, Species, SEAc)

Mean_position <- read.csv(file.path(here(), "exported_data", 
                                    "Meanposition_SEAc.csv"), header = T)
names(Mean_position)[3] <- "Code"
```

```{r merge datasets}

Pop_metrics <- merge(Niche_metrics, Mean_position, 
                     by = c("Pond", "Code", "Species"))
dim(Pop_metrics)[1] == dim(Niche_metrics)[1]
dim(Pop_metrics)[1] == dim(Mean_position)[1]

for (i in 1:length(Pop_metrics$Pond)) {
  if (Pop_metrics$Species[i] %in% c("Agile frog", "Tree frog")) {
    Pop_metrics$Group[i] = "Tadpoles"
    Pop_metrics$Clade[i] = "Amphibians"
  } else if (Pop_metrics$Species[i] %in% c("Palmate newt", "Marbled newt")) {
    Pop_metrics$Group[i] = "Newts"
    Pop_metrics$Clade[i] = "Amphibians"
  } else {
    Pop_metrics$Group[i] = "Crayfish"
    Pop_metrics$Clade[i] = "Crayfish"
  }
}

#Create a dataset only for amphibian populations
Pop_amphib <- Pop_metrics %>% filter(Clade == "Amphibians")

#Merge metrics and explicative variables
Pop_dataset <- merge(Pop_amphib, Envt, by = "Pond")
print("Are there still 41 populations in the dataset?")
dim(Pop_dataset)[1] == 41

```



# 2. Models: variation in amphibian niche metrics

For each metric tested, we (1) check GOF of a simple model, and transform the metric if necessary, (2) run the model selection and (3) plot the figure corresponding to the selected models.

### SEAc

```{r check GOF for SEAc, include=FALSE}
mod1 <- lmer(SEAc ~ scale(Canopycover)+ (1|Species), 
             na.action = "na.fail", data = Pop_dataset)
plotresid(mod1)
summary(mod1)

mod1 <- lm(SEAc ~ scale(Canopycover), 
             na.action = "na.fail", data = Pop_dataset)
plotresid(mod1)

```

Residuals are not ok, SEAc must be transformed.
The random species effect has a very low variance and must be discarded

```{r check GOF for log-transformed SEAc, include=FALSE}

mod2 <- lm(log(SEAc) ~ scale(Canopycover),
             na.action = "na.fail", data = Pop_dataset)
plotresid(mod2)
```

Residuals are greatly improved, we go with the transformed SEAc.

```{r dredge SEAc model, include=FALSE}

# Test effect of candidate variables
mod_SEA <- lm(log(SEAc) ~ scale(Area) + 
                  scale(Canopycover) + 
                  scale(Aq_VegCover) +
                  scale(CrayfishCPUE) +
                as.factor(Crayfish01) +
                  scale(Agile_frog) + 
                  scale(Tree_frog) + 
                  scale(Palmate_newt) + 
                  scale(Marbled_newtCPUE) +
                  Group, 
                na.action = "na.fail", 
                data = Pop_dataset)

SEA_dredge <- dredge(mod_SEA,
                     rank = AICc, m.lim = c(0, 2), extra="R^2")
SEA_select <- subset(SEA_dredge, delta < 2)
SEA_select

write.csv(SEA_dredge, file.path(here(), "exported_data", 
                             "SEAc_selection.csv"), row.names = F)

```

One model is retained for SEAc.

 
```{r check the content of each model for SEAc, include=FALSE}
SEA_dredge <- dredge(mod_SEA, rank = AICc, 
                     m.lim = c(0,2), extra="R^2", evaluate=FALSE)

SEA_dredge$'67'
Model_selected1 <- lm(log(SEAc) ~ scale(Agile_frog) + scale(Marbled_newtCPUE),
             na.action="na.fail", data = Pop_dataset)
plotresid(Model_selected1)

summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)
Anova(Model_selected1)


SEA_dredge$'49'
Model_selected1 <- lm(log(SEAc) ~ scale(Canopycover) +
                          scale(CrayfishCPUE),
             na.action="na.fail", data = Pop_dataset)
plotresid(Model_selected1)

# summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)
Anova(Model_selected1)



SEA_dredge$'6'
Model_selected1 <- lm(log(SEAc) ~ scale(Agile_frog) +
                          as.factor(Crayfish01),
             na.action="na.fail", data = Pop_dataset)
plotresid(Model_selected1)

# summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)
Anova(Model_selected1)
```

# Figure SEAc
```{r corresponding figures for SEAc, include=FALSE}

library(cowplot)

Pop_dataset$Species <- factor(Pop_dataset$Species, levels = c("Crayfish",
                                                    "Agile frog",
                                                    "Tree frog",
                                                    "Palmate newt",
                                                    "Marbled newt"))

SEAc_marbre <- ggplot(Pop_dataset, aes(y = log(SEAc), 
                                       x = scale(Marbled_newtCPUE)
                                       # , group = Species
                                       )) +
  geom_smooth(method="lm",se=F, col = "black") +
    geom_point(aes(col = Species), size = 2, show.legend = F) +
    scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Marbled newt abundance, scaled") +
  ylab("log(SEAc)") +
  theme_classic() +
  theme(legend.position="none")

SEAc_marbre


SEAc_agile <- ggplot(Pop_dataset,
                     aes(y = log(SEAc),
                         x = scale(Agile_frog)
                         # , group = Species
                         )) +
  geom_smooth(method="lm",se=F, col = "black") +
    geom_point(aes(col = Species), size = 2, show.legend = F) +
    scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Agile frog density, scaled") +
  ylab(NULL) +
  theme_classic() +
  theme(legend.position="none")

SEAc_agile


Pop_dataset %<>% mutate(Crayfish01 = dplyr::recode(Crayfish01, 
                                          "1" = "present",
                                          "0" = "absent"))
SEAc_crayfish <- ggplot(Pop_dataset, 
             aes(y = log(SEAc), 
                 x = as.factor(Crayfish01)
                 # , group = Species
                 )) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
  scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
    xlab("Crayfish") +
  ylab(NULL) +
  theme_classic() +
  theme(legend.position="none")

SEAc_crayfish


SEAc <- grid.arrange(SEAc_marbre, SEAc_agile, SEAc_crayfish, ncol = 3)
```


### Mean TP

```{r check GOF for mean TP, include=FALSE}

# Check residuals

mod1 <- lmer(TPavg ~ scale(Canopycover) + (1|Species), na.action="na.fail", data=Pop_dataset)
plotresid(mod1) #residuals ok

mod1 <- lm(TPavg ~ scale(Canopycover), na.action="na.fail", data=Pop_dataset)
plotresid(mod1) #residuals ok

```

Again, very low variance for the random effect species, so we remove it.
 
```{r dredge TP model, include=FALSE}

# Test effect of candidate variables
mod_TP <- lm(TPavg ~ scale(Area) + scale(Canopycover) + 
                 scale(Aq_VegCover) +
                 scale(CrayfishCPUE) + as.factor(Crayfish01) + scale(Agile_frog) + 
                 scale(Tree_frog) + scale(Palmate_newt) +
                 scale(Marbled_newtCPUE) +
                 Group, 
               na.action="na.fail", data=Pop_dataset)

TP_dredge <- dredge(mod_TP, rank = AICc, m.lim = c(0,2), extra="R^2")
subset(TP_dredge, delta < 2)

write.csv(TP_dredge, file.path(here(), "exported_data", 
                             "TP_selection.csv"), row.names = F)
```

 
```{r check the content of each model for TP, include=FALSE}

TP_dredge <- dredge(mod_TP, rank = AICc, m.lim = c(0,2), extra="R^2", evaluate=FALSE)

TP_dredge$'18'
Model_selected1 <- lm(TPavg ~ scale(Canopycover) + Group, 
                         na.action="na.fail", data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)


TP_dredge$'34'
Model_selected1 <- lm(TPavg ~ scale(CrayfishCPUE) + Group, 
                        na.action="na.fail", 
                        data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)


Model_selected1 <- lm(TPavg ~ as.factor(Crayfish01) + Group, 
                        na.action="na.fail", 
                        data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)

```

 
```{r corresponding figures for TP, include=FALSE}

TP_canopy <- ggplot(Pop_dataset, aes(y = TPavg, x = scale(Canopycover))) +
  geom_smooth(method="lm",se=F, col = "black") +
    geom_point(aes(col = Species), size = 2) +
    scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Canopy cover, scaled") +
  ylab(NULL) +
  theme_classic() + 
  theme(legend.position="none")

  TP_canopy

  
  TP_crayfish <- ggplot(Pop_dataset, aes(y = TPavg, x = as.factor(Crayfish01))) +
    geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
    scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Crayfish") +
      ylab(NULL) +
  theme_classic() +
  theme(legend.position="none")

  TP_crayfish


Pop_dataset$Group <- factor(Pop_dataset$Group, levels = c("Tadpoles", "Newts")) 
  
    TP_group <- ggplot(Pop_dataset, aes(y = TPavg, x = Group)) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Group") +
  ylab("Mean TP") + 
  theme_classic() +
  theme(legend.position="none")

  TP_group

  
TP <- grid.arrange(TP_group, TP_canopy, TP_crayfish, ncol = 3)
  
```

 
### Mean d13C
 
```{r check GOF for D13C, include=FALSE}
 
# Check residuals

mod1 <- lmer(D13Cavg ~ scale(Canopycover) + (1|Species), 
             na.action="na.fail", data=Pop_dataset)
plotresid(mod1) #residuals ok


mod1 <- lm(D13Cavg ~ scale(Palmate_newt) + Group, 
             na.action="na.fail", data=Pop_dataset)
plotresid(mod1) #residuals ok
```

The random effect species is again very low, we remove it.

```{r dredge D13C model, include=FALSE}
 
# Test effect of candidate variables
mod_D13C <- lm(D13Cavg ~ scale(Area) + scale(Canopycover) + 
                   scale(Aq_VegCover) +
                   scale(CrayfishCPUE) + as.factor(Crayfish01) +
                   scale(Agile_frog) + scale(Tree_frog) + 
                   scale(Palmate_newt) + scale(Marbled_newtCPUE) +
                   Group, 
                 na.action="na.fail", data=Pop_dataset)

D13C_dredge <- dredge(mod_D13C, rank = AICc, m.lim = c(0,2), extra="R^2")
subset(D13C_dredge, delta < 2)

write.csv(D13C_dredge, file.path(here(), "exported_data", 
                             "D13C_selection.csv"), row.names = F)

```

 
```{r check the content of each model for D13C, include=FALSE}

D13C_dredge <- dredge(mod_D13C, rank = AICc, m.lim = c(0,2), extra="R^2", evaluate=FALSE)

D13C_dredge$'130'
Model_selected1 <- lm(D13Cavg ~ scale(Palmate_newt) + Group, 
                      na.action="na.fail", data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)


D13C_dredge$'34'
Model_selected1 <- lm(D13Cavg ~ scale(CrayfishCPUE) + Group, 
                      na.action="na.fail", data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)


Model_selected1 <- lm(D13Cavg ~ as.factor(Crayfish01) + Group, 
                      na.action="na.fail", data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)
```

 
```{r corresponding figures for D13C, include=FALSE}

D13C_group <- ggplot(Pop_dataset, aes(y = D13Cavg, x = Group)) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Group") +
  ylab(bquote('Mean '*delta^13~C["cor"]*'')) +
  theme_classic() +
  theme(legend.position="none")
D13C_group


D13C_palmate <- ggplot(Pop_dataset, aes(y = D13Cavg, 
                              x = scale(Palmate_newt))) +
  geom_point(aes(col = Species), size = 2, show.legend = F) +
  geom_smooth(method="lm",se=F, col = "black") +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
    ylab(NULL) +
  xlab("Palmate newt abundance, scaled") +
  theme_classic()
D13C_palmate


D13C_crayfish <- ggplot(Pop_dataset, aes(y = D13Cavg, 
                              x = as.factor(Crayfish01))) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2, show.legend = F) +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
    ylab(NULL) +
  xlab("Crayfish") +
  theme_classic()
D13C_crayfish

D13C <- grid.arrange(D13C_group, D13C_palmate, D13C_crayfish, ncol = 3)

```


### Figure 3
```{r figure 3}

# Get the legend
legend <- ggplot(Pop_dataset, aes(y = D13Cavg, 
                              x = as.factor(Crayfish01))) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
    ylab(NULL) +
  xlab("Crayfish") +
  theme_classic() +
  theme(legend.position = "bottom")

library(cowplot)
get_legend(legend)

g <- ggdraw() +
  draw_plot(SEAc, x = 0, y = 0.7, width = 1, height = 0.3) +
  draw_plot(TP, 0, 0.4, 1, .3) +
  draw_plot(D13C, 0, 0.1, 1, 0.3) +
  draw_plot(get_legend(legend), 0, 0, 1, 0.1) +
  draw_plot_label(c("A", "B", "C"), 
                  c(0, 0, 0), 
                  c(1, 0.7, 0.4))

ggsave("./figures/Figure3c.jpg", g, width = 8, height = 8)
```


