---
title: "#6: Statistical models"
author: 
- Nadege Belouard^[UMR CNRS EcoBio, UMR DECOD, nadege.belouard@gmail.com]
date: "23/03/2023"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2
  html_document:
    toc: TRUE
    toc_depth: 3
params:
  show_code: FALSE
  export_figures: TRUE
editor_options: 
  chunk_output_type: console
---

This vignette tests the link between the isotopic niche metrics of amphibians and candidate variables, including crayfish.


# 1. Setup

We begin by loading the required libraries and datasets.

```{r setup, include = FALSE}

#Load packages
library(lme4)
library(car)
library(MuMIn)
library(ggplot2)
library(RVAideMemoire)
library(ggrepel)
library(magrittr)
library(Hmisc)
library(usdm)
library(tidyr)
library(dplyr)
library(tibble)
library(gridExtra)
library(here)
library(cowplot)

```

Then we load the datasets: environmental variables, population metrics and mean position.

```{r load datasets}
#Load datasets
Envt <- read.csv(here::here("data_raw", "Raw_environment.csv"), h = T)
Envt %<>% mutate(Crayfish01 = ifelse(CrayfishCPUE == 0, 0, 1))
         
# Dataset
Niche_metrics <- read.csv(file.path(here(), "exported_data", 
                                    "Population_metrics.csv"), h = T) %>% 
  dplyr::select(Code, Pond, Species, SEAc)

Mean_position <- read.csv(file.path(here(), "exported_data", 
                                    "Meanposition_SEAc.csv"), header = T)
names(Mean_position)[3] <- "Code"
```

We merge the different datasets and create a "Group" column (newts or tadpoles). We also create a dataset that keeps only amphibian populations for analysis of the variation in niche metrics.
```{r merge datasets}

Pop_metrics <- merge(Niche_metrics, Mean_position, 
                     by = c("Pond", "Code", "Species"))
dim(Pop_metrics)[1] == dim(Niche_metrics)[1]
dim(Pop_metrics)[1] == dim(Mean_position)[1]

for (i in 1:length(Pop_metrics$Pond)) {
  if (Pop_metrics$Species[i] %in% c("Agile frog", "Tree frog")) {
    Pop_metrics$Group[i] = "Tadpoles"
    Pop_metrics$Clade[i] = "Amphibians"
  } else if (Pop_metrics$Species[i] %in% c("Palmate newt", "Marbled newt")) {
    Pop_metrics$Group[i] = "Newts"
    Pop_metrics$Clade[i] = "Amphibians"
  } else {
    Pop_metrics$Group[i] = "Crayfish"
    Pop_metrics$Clade[i] = "Crayfish"
  }
}

#Create a dataset only for amphibian populations
Pop_amphib <- Pop_metrics %>% filter(Clade == "Amphibians")

#Merge metrics and explicative variables
Pop_dataset <- merge(Pop_amphib, Envt, by = "Pond")
print("Are there still 41 populations in the dataset?")
dim(Pop_dataset)[1] == 41

```



# 2. Models: variation in amphibian niche metrics

For each metric tested, we (1) check GOF of a simple model, and transform the metric if necessary, (2) run the model selection and (3) plot the figure corresponding to the selected models.

### SEAc

```{r check GOF for SEAc, include=FALSE}
mod1 <- lmer(SEAc ~ scale(Canopycover)+ (1|Species), 
             na.action = "na.fail", data = Pop_dataset)
plotresid(mod1)
summary(mod1)

mod1 <- lm(SEAc ~ scale(Canopycover), 
             na.action = "na.fail", data = Pop_dataset)
plotresid(mod1)

```

Residuals are not ok, SEAc must be transformed.
The random species effect has a very low variance and must be discarded

```{r check GOF for log-transformed SEAc, include=FALSE}

mod2 <- lm(log(SEAc) ~ scale(Canopycover),
             na.action = "na.fail", data = Pop_dataset)
plotresid(mod2)
```

Residuals are greatly improved, we go with the transformed SEAc.

```{r dredge SEAc model, include=FALSE}

# Test effect of candidate variables
mod_SEA <- lm(log(SEAc) ~ scale(Area) + 
                  scale(Canopycover) + 
                  scale(Aq_VegCover) +
                  scale(CrayfishCPUE) +
                as.factor(Crayfish01) +
                  scale(Agile_frog) + 
                  scale(Tree_frog) + 
                  scale(Palmate_newt) + 
                  scale(Marbled_newtCPUE) +
                  Group, 
                na.action = "na.fail", 
                data = Pop_dataset)

SEA_dredge <- dredge(mod_SEA,
                     rank = AICc, m.lim = c(0, 2), extra="R^2")
SEA_select <- subset(SEA_dredge, delta < 2)
SEA_select

write.csv(SEA_dredge, file.path(here(), "exported_data", 
                             "SEAc_selection.csv"), row.names = F)

```

One model is retained for SEAc, it contains the variables agile frog and marbled newt.

We check the content of this model as well as the first model containing either Crayfish abundance or Crayfish presence.
```{r check the content of each model for SEAc, include=FALSE}
SEA_dredge <- dredge(mod_SEA, rank = AICc, 
                     m.lim = c(0,2), extra="R^2", evaluate=FALSE)

# First model
SEA_dredge$'67'
Model_selected1 <- lm(log(SEAc) ~ scale(Agile_frog) + scale(Marbled_newtCPUE),
             na.action="na.fail", data = Pop_dataset)
plotresid(Model_selected1)

summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)
Anova(Model_selected1)


# First model with crayfish abundance
SEA_dredge$'49'
Model_selected1 <- lm(log(SEAc) ~ scale(Canopycover) +
                          scale(CrayfishCPUE),
             na.action="na.fail", data = Pop_dataset)
plotresid(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)
Anova(Model_selected1)


#First model with crayfish presence
SEA_dredge$'6'
Model_selected1 <- lm(log(SEAc) ~ scale(Agile_frog) +
                          as.factor(Crayfish01),
             na.action="na.fail", data = Pop_dataset)
plotresid(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)
Anova(Model_selected1)
```

### Figure with SEAc
```{r corresponding figures for SEAc, include=FALSE}

# Organize factor levels
Pop_dataset$Species <- factor(Pop_dataset$Species, levels = c("Crayfish",
                                                    "Agile frog",
                                                    "Tree frog",
                                                    "Palmate newt",
                                                    "Marbled newt"))

Pop_dataset %<>% mutate(Crayfish01 = dplyr::recode(Crayfish01, 
                                          "1" = "present",
                                          "0" = "absent"))


# Figure with marbled newt
SEAc_marbre <- ggplot(Pop_dataset, aes(y = log(SEAc), 
                                       x = scale(Marbled_newtCPUE)
                                       # , group = Species
                                       )) +
  geom_smooth(method="lm",se=F, col = "black") +
    geom_point(aes(col = Species), size = 2, show.legend = F) +
    scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Marbled newt abundance, scaled") +
  ylab("log(SEAc)") +
  theme_classic() +
  theme(legend.position="none")

SEAc_marbre

# Figure with agile frog
SEAc_agile <- ggplot(Pop_dataset,
                     aes(y = log(SEAc),
                         x = scale(Agile_frog)
                         # , group = Species
                         )) +
  geom_smooth(method="lm",se=F, col = "black") +
    geom_point(aes(col = Species), size = 2, show.legend = F) +
    scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Agile frog density, scaled") +
  ylab(NULL) +
  theme_classic() +
  theme(legend.position="none")

SEAc_agile


# Figure with crayfish presence
SEAc_crayfish <- ggplot(Pop_dataset, 
             aes(y = log(SEAc), 
                 x = as.factor(Crayfish01)
                 # , group = Species
                 )) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
  scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
    xlab("Crayfish") +
  ylab(NULL) +
  theme_classic() +
  theme(legend.position="none")

SEAc_crayfish

# Arrange the 3 figures on a row
SEAc <- grid.arrange(SEAc_marbre, SEAc_agile, SEAc_crayfish, ncol = 3)
```


## Mean TP

First check the model GOF
```{r check GOF for mean TP, include=FALSE}

# Check residuals

mod1 <- lmer(TPavg ~ scale(Canopycover) + (1|Species), na.action="na.fail", data=Pop_dataset)
plotresid(mod1) #residuals ok

mod1 <- lm(TPavg ~ scale(Canopycover), na.action="na.fail", data=Pop_dataset)
plotresid(mod1) #residuals ok

```

Again, very low variance for the random effect species, so we remove it. Residuals look normal.
 
```{r dredge TP model, include=FALSE}

# Test effect of candidate variables
mod_TP <- lm(TPavg ~ scale(Area) + scale(Canopycover) + 
                 scale(Aq_VegCover) +
                 scale(CrayfishCPUE) + as.factor(Crayfish01) + scale(Agile_frog) + 
                 scale(Tree_frog) + scale(Palmate_newt) +
                 scale(Marbled_newtCPUE) +
                 Group, 
               na.action="na.fail", data=Pop_dataset)

TP_dredge <- dredge(mod_TP, rank = AICc, m.lim = c(0,2), extra="R^2")
subset(TP_dredge, delta < 2)

write.csv(TP_dredge, file.path(here(), "exported_data", 
                             "TP_selection.csv"), row.names = F)
```

A single model is kept by the selection, it contains the variables group and canopy cover. 
We check the content of this model as well as the first model containing either Crayfish abundance or Crayfish presence.
```{r check the content of each model for TP, include=FALSE}

TP_dredge <- dredge(mod_TP, rank = AICc, m.lim = c(0,2), extra="R^2", evaluate=FALSE)

# First model
TP_dredge$'18'
Model_selected1 <- lm(TPavg ~ scale(Canopycover) + Group, 
                         na.action="na.fail", data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)

#First model with crayfish abundance
TP_dredge$'34'
Model_selected1 <- lm(TPavg ~ scale(CrayfishCPUE) + Group, 
                        na.action="na.fail", 
                        data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)

#First model with crayfish presence
Model_selected1 <- lm(TPavg ~ as.factor(Crayfish01) + Group, 
                        na.action="na.fail", 
                        data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)

```

 
### Figure with TP
```{r corresponding figures for TP, include=FALSE}

# Figure with group
Pop_dataset$Group <- factor(Pop_dataset$Group, levels = c("Tadpoles", "Newts")) 
  
    TP_group <- ggplot(Pop_dataset, aes(y = TPavg, x = Group)) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Group") +
  ylab("Mean TP") + 
  theme_classic() +
  theme(legend.position="none")

  TP_group

  
# Figure with canopy cover
TP_canopy <- ggplot(Pop_dataset, aes(y = TPavg, x = scale(Canopycover))) +
  geom_smooth(method="lm",se=F, col = "black") +
    geom_point(aes(col = Species), size = 2) +
    scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Canopy cover, scaled") +
  ylab(NULL) +
  theme_classic() + 
  theme(legend.position="none")

  TP_canopy

  
# Figure with crayfish presence
  TP_crayfish <- ggplot(Pop_dataset, aes(y = TPavg, x = as.factor(Crayfish01))) +
    geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
    scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Crayfish") +
      ylab(NULL) +
  theme_classic() +
  theme(legend.position="none")

  TP_crayfish



# We put all three figures on a single row
TP <- grid.arrange(TP_group, TP_canopy, TP_crayfish, ncol = 3)
  
```

 
## Mean d13C

We check the GOF of the model with D13C
```{r check GOF for D13C, include=FALSE}

mod1 <- lmer(D13Cavg ~ scale(Canopycover) + (1|Species), 
             na.action="na.fail", data=Pop_dataset)
plotresid(mod1) #residuals ok


mod1 <- lm(D13Cavg ~ scale(Palmate_newt) + Group, 
             na.action="na.fail", data=Pop_dataset)
plotresid(mod1) #residuals ok
```

The random effect species is again very low, we remove it. Very few points don't follow normality of the residuals. Transformations are however inappropriate for this variable because of negative values. We keep the variable untransformed.

```{r dredge D13C model, include=FALSE}
 
# Test effect of candidate variables
mod_D13C <- lm(D13Cavg ~ scale(Area) + scale(Canopycover) + 
                   scale(Aq_VegCover) +
                   scale(CrayfishCPUE) + as.factor(Crayfish01) +
                   scale(Agile_frog) + scale(Tree_frog) + 
                   scale(Palmate_newt) + scale(Marbled_newtCPUE) +
                   Group, 
                 na.action="na.fail", data=Pop_dataset)

D13C_dredge <- dredge(mod_D13C, rank = AICc, m.lim = c(0,2), extra="R^2")
subset(D13C_dredge, delta < 2)

write.csv(D13C_dredge, file.path(here(), "exported_data", 
                             "D13C_selection.csv"), row.names = F)

```

Again, a single model is kept, it contains the variable gruop and palmate newt.
We check the content of this model as well as the first model containing either Crayfish abundance or Crayfish presence.
 
```{r check the content of each model for D13C, include=FALSE}

D13C_dredge <- dredge(mod_D13C, rank = AICc, m.lim = c(0,2), extra="R^2", evaluate=FALSE)

#First model
D13C_dredge$'130'
Model_selected1 <- lm(D13Cavg ~ scale(Palmate_newt) + Group, 
                      na.action="na.fail", data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)

# First model with crayfish abundance
D13C_dredge$'34'
Model_selected1 <- lm(D13Cavg ~ scale(CrayfishCPUE) + Group, 
                      na.action="na.fail", data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)

# First model with crayfish presence
Model_selected1 <- lm(D13Cavg ~ as.factor(Crayfish01) + Group, 
                      na.action="na.fail", data=Pop_dataset)
plotresid(Model_selected1)
summary(Model_selected1)
res <- summary(Model_selected1)
round(res$coefficients,2)
```

### Figure with D13C 
```{r corresponding figures for D13C, include=FALSE}

#Figure with group
D13C_group <- ggplot(Pop_dataset, aes(y = D13Cavg, x = Group)) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
  xlab("Group") +
  ylab(bquote('Mean '*delta^13~C["cor"]*'')) +
  theme_classic() +
  theme(legend.position="none")
D13C_group

#Figure with palmate newt
D13C_palmate <- ggplot(Pop_dataset, aes(y = D13Cavg, 
                              x = scale(Palmate_newt))) +
  geom_point(aes(col = Species), size = 2, show.legend = F) +
  geom_smooth(method="lm",se=F, col = "black") +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
    ylab(NULL) +
  xlab("Palmate newt abundance, scaled") +
  theme_classic()
D13C_palmate

# Figure with crayfish presence
D13C_crayfish <- ggplot(Pop_dataset, aes(y = D13Cavg, 
                              x = as.factor(Crayfish01))) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2, show.legend = F) +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
    ylab(NULL) +
  xlab("Crayfish") +
  theme_classic()
D13C_crayfish

# Put all 3 figures on a single row
D13C <- grid.arrange(D13C_group, D13C_palmate, D13C_crayfish, ncol = 3)

```


## Figure 3
Finally we assemble all the different metrics on a single figure
```{r figure 3}

# Get the legend
legend <- ggplot(Pop_dataset, aes(y = D13Cavg, 
                              x = as.factor(Crayfish01))) +
  geom_boxplot(show.legend = F) +
      geom_jitter(aes(col = Species), size = 2) +
scale_color_manual(values = c("#0072B2", "#009E73", 
                                "#56B4E9", "#CC79A7")) +
    ylab(NULL) +
  xlab("Crayfish") +
  theme_classic() +
  theme(legend.position = "bottom")

get_legend(legend)

g <- ggdraw() +
  draw_plot(SEAc, x = 0, y = 0.7, width = 1, height = 0.3) +
  draw_plot(TP, 0, 0.4, 1, .3) +
  draw_plot(D13C, 0, 0.1, 1, 0.3) +
  draw_plot(get_legend(legend), 0, 0, 1, 0.1) +
  draw_plot_label(c("A", "B", "C"), 
                  c(0, 0, 0), 
                  c(1, 0.7, 0.4))

ggsave("./figures/Figure3c.jpg", g, width = 8, height = 8)
```


