---
title: "Coexistence crayfish/amphibians #6: Niche comparison"
author: 
- Nadege Belouard^[UMR CNRS EcoBio, UMR INRAE DECOD, nadege.belouard@gmail.com]
date: "10/08/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2
  html_document:
    toc: TRUE
    toc_depth: 3
params:
  show_code: FALSE
  export_figures: TRUE
editor_options: 
  chunk_output_type: console
---

This vignette compares niche metrics between the different study species in ponds

# 1. Setup

We begin by loading the required libraries and datasets.

```{r setup, warning = FALSE, message = FALSE}
library(tidyr)
library(dplyr)
library(magrittr)
library(reshape2)
library(here)
library(PairedData)
```

```{r load datasets per pop}
Niche_metrics <- read.csv(file.path(here(), "exported_data", "Population_metrics_complete.csv"), h = T)
Mean_position <- read.csv(file.path(here(), "exported_data", "Meanposition_SEAc_complete.csv"), header = T)
```

We merge both datasets (one contains the Layman metrics and the other the SEAc).

```{r merge datasets}
# Rename a column in one dataset so it is common with the other dataset
Mean_position %<>% rename(Code = Pop)
Pop_metrics <- merge(Niche_metrics, Mean_position, by = c("Pond", "Code"))
```


# 2. Compare crayfish and amphibians niches

## A. Agile frog/All crayfish (n = 9)
Prepare the dataset and show mean values
```{r prep crayfish and agile frog dataset}
metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish") 
metrics_agile <- Pop_metrics %>% filter(Species == "Agile frog") 

metrics <- rbind(metrics_crayfish, metrics_agile)
metrics %<>% filter(Pond %in% c(1016, 111, 112, 66, 72, 888, 89, 91, 949))

metrics %>% group_by(TaxaUsed) %>% summarise(
                                             meanSEAc = mean(SEAc),
                                             meanNR = mean(NRcp)
                                             )
```

Wilcoxon tests
```{r wilcoxon tests agile frog crayfish}

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 1, p-value = 0.007812 **

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 5, p-value = 0.03906 *

```

Paired permutation test 
```{r permut tests agile frog/crayfish}
metrics %<>% dplyr::select(Pond, NRcp, SEAc, TaxaUsed)
metrics_pivoted <- pivot_wider(metrics, names_from = TaxaUsed, values_from = c(NRcp, SEAc))

# Test par permutation sur la différence de moyenne:
# Calculate the difference between the two species
metrics_pivoted %<>% mutate(
                            diffNR = NRcp_Crayfish - `NRcp_Agile frog`,
                            diffSEAc = SEAc_Crayfish - `SEAc_Agile frog`
                            )


# permutation test: set the number of permutations 
sim <- 10^4 #number of permutations

#create an empty vector to store the simulated averages
permNR <- rep(NA, sim) #
permSEAc <- rep(NA, sim) #

k <- nrow(metrics_pivoted) # number of individuals to sample/permute
set.seed(2022) # make sure the results are going to be the same every time we run the code

# permute data
for (i in 1:sim) { #in each simulation
  # randomly select some pairs to have their difference multiplied by negative to indicate the other species should have the highest size
tirage = sample(c(-1,1), size = k, replace = T)
permSEAc[i] = mean(metrics_pivoted$diffSEAc * tirage)
permNR[i] = mean(metrics_pivoted$diffNR * tirage)
}

#add the observed difference
permNR[10000] = mean(metrics_pivoted$diffNR)
permSEAc[10000] = mean(metrics_pivoted$diffSEAc)


# Table disjonctive qui compte le nombre de valeurs dans perm qui sont superieures ou egales
# a la difference observee entre les moyennes, puis calcule la valeur de p

tablepropNR = table(abs(permNR) >= abs(mean(metrics_pivoted$diffNR)))
pNR = tablepropNR[2]/10000
pNR #p =0.0419

tablepropSEAc = table(abs(permSEAc) >= abs(mean(metrics_pivoted$diffSEAc)))
pSEAc = tablepropSEAc[2]/10000
pSEAc # p = 0.008

```


## B. Tree frog/All crayfish (n = 4)
Prepare the dataset and show mean values
```{r test differences all crayfish and tree frog}
metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish")
metrics_treefrog <- Pop_metrics %>% filter(TaxaUsed == "Tree frog") 

metrics <- rbind(metrics_crayfish, metrics_treefrog)
metrics %<>% filter(Pond %in% c(66, 89, 91, 949))

metrics %>% group_by(TaxaUsed) %>% summarise(
                                             meanSEAc = mean(SEAc),
                                             meanNR = mean(NRcp)
                                             )
```

Wilcoxon tests
```{r wilcoxon tests tree frog crayfish}

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
#V = 2, p-value = 0.375

```


Permutation paired t-test
```{r permut tests tree frog crayfish}
metrics %<>% dplyr::select(Pond, NRcp, SEAc, TaxaUsed)
metrics_pivoted <- pivot_wider(metrics, names_from = TaxaUsed, values_from = c(NRcp, SEAc))


# Test par permutation sur la différence de moyenne: TAcp
# Calculate the difference
metrics_pivoted %<>% mutate(
                            diffNR = NRcp_Crayfish - `NRcp_Tree frog`,
                            diffSEAc = SEAc_Crayfish - `SEAc_Tree frog`
                            )


# permutation test: set the number of permutations 
sim <- 10^4 #number of permutations
permNR <- rep(NA, sim) #
permSEAc <- rep(NA, sim) #


k <- nrow(metrics_pivoted)# number of individuals to sample/permute
set.seed(2022) # make sure the results are going to be the same

# permute data
for (i in 1:sim) {
  # randomly select some pairs to have their difference multiplied by negative to indicate the other species should have the highest size
tirage = sample(c(-1,1), size = k, replace = T)
permSEAc[i] = mean(metrics_pivoted$diffSEAc * tirage)
permNR[i] = mean(metrics_pivoted$diffNR * tirage)
}

permNR[10000] = mean(metrics_pivoted$diffNR)
permSEAc[10000] = mean(metrics_pivoted$diffSEAc)
# permSDNND[10000] = mean(metrics_pivoted$diffSDNND)

# Table disjonctive qui compte le nombre de valeurs dans perm qui sont superieures ou egales
# a la difference observee entre les moyennes

tablepropNR = table(abs(permNR) >= abs(mean(metrics_pivoted$diffNR)))
pNR = tablepropNR[2]/10000
pNR #p = 0.3816 no diff

tablepropSEAc = table(abs(permSEAc) >= abs(mean(metrics_pivoted$diffSEAc)))

pSEAc = tablepropSEAc[2]/10000
pSEAc 
tablepropSEAc
# All simulations are TRUE, so p > 0.9999, no difference

```


## C. Palmate newt/All crayfish (n = 5)
Prepare the dataset and show mean values
```{r test differences crayfish adult and palmate newt}

metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish") 
metrics_palmate <- Pop_metrics %>% filter(TaxaUsed == "Palmate newt")

metrics <- rbind(metrics_crayfish, metrics_palmate)
metrics %<>% filter(Pond %in% c(111, 112, 66, 91, 949))

metrics %>% group_by(TaxaUsed) %>% summarise(
                                             meanSEAc = round(mean(SEAc),2),
                                             meanNR = round(mean(NRcp),2)
                                             )
```

Wilcoxon tests
```{r wilcoxon tests palmate newt crayfish}

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625 .

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 10, p-value = 0.625

```


Permutation paired t-test: 
```{r permut tests palmate newt crayfish}
metrics %<>% dplyr::select(Pond, NRcp, SEAc, TaxaUsed)
metrics_pivoted <- pivot_wider(metrics, names_from = TaxaUsed, values_from = c(NRcp, SEAc))

# Test par permutation sur la différence de moyenne: 
# Calculate the difference
metrics_pivoted %<>% mutate(
                            diffNR = NRcp_Crayfish - `NRcp_Palmate newt`,
                            diffSEAc = SEAc_Crayfish - `SEAc_Palmate newt`
                            )


# permutation test: set the number of permutations 
sim <- 10^4 #number of permutations
permNR <- rep(NA, sim) #
permSEAc <- rep(NA, sim) #


k <- nrow(metrics_pivoted)# number of individuals to sample/permute
set.seed(2022) # make sure the results are going to be the same

# permute data
for (i in 1:sim) {
  # randomly select some pairs to have their difference multiplied by negative to indicate the other species should have the highest size
tirage = sample(c(-1,1), size = k, replace = T)
permSEAc[i] = mean(metrics_pivoted$diffSEAc * tirage)
permNR[i] = mean(metrics_pivoted$diffNR * tirage)
}

permNR[10000] = mean(metrics_pivoted$diffNR)
permSEAc[10000] = mean(metrics_pivoted$diffSEAc)


# Table disjonctive qui compte le nombre de valeurs dans perm qui sont superieures ou egales
# a la difference observee entre les moyennes

tablepropNR = table(abs(permNR) >= abs(mean(metrics_pivoted$diffNR)))
pNR = tablepropNR[2]/10000
pNR # p = 0.4966

tablepropSEAc = table(abs(permSEAc) >= abs(mean(metrics_pivoted$diffSEAc)))
pSEAc = tablepropSEAc[2]/10000
pSEAc # p = 0.0652

```



## D. Marbled newt/All crayfish (n = 0)
No stats


