---
title: "Coexistence crayfish/amphibians #6: Niche comparison"
author: 
- Nadege Belouard^[UMR CNRS EcoBio, UMR INRAE DECOD, nadege.belouard@gmail.com]
date: "10/08/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 3
  html_document:
    toc: FALSE
    toc_depth: 3
params:
  show_code: FALSE
  export_figures: TRUE
editor_options: 
  chunk_output_type: console
---



# 1. Setup

This vignette compares niche metrics between the different study species in ponds.
Only SEAc and NR are compared because all metrics are extremely correlated to them.
CDreg (the distance between the centroid of each population and the centroid of the regional niche of the species) is added even though it is also very correlated to the other metrics, for exploratory purposes.

We begin by loading the required libraries and datasets.

```{r setup, warning = FALSE, message = FALSE}
library(tidyr)
library(dplyr)
library(magrittr)
library(reshape2)
library(here)
library(PairedData)
```

```{r load datasets per pop}
Niche_metrics <- read.csv(file.path(here(), "data", "Population_metrics_complete.csv"), h = T)
Mean_position <- read.csv(file.path(here(), "data", "Meanposition_SEAc_complete.csv"), header = T)
CDreg <- read.csv(file.path(here(), "exported_data", "CD_global.csv"), header = T)
```

We merge datasets (one contains the Layman metrics, one the SEAc, the last one CDregional).

```{r merge datasets}
# Rename a column in one dataset so it is common with the other dataset
Mean_position %<>% rename(Code = Pop)
Pop_metrics <- merge(Niche_metrics, Mean_position, by = c("Pond", "Code"))

# select only the columns we are interested in
CDreg %<>% dplyr::select(Code, Species, CDcp) %>% rename(Pond = Code, CDreg = CDcp)  
Pop_metrics <- merge(Pop_metrics, CDreg, by = c("Pond", "Species"))
```


# 2. Compare crayfish and amphibians niches

## A. Agile frog/All crayfish (n = 9)
### Prepare the dataset and show mean values
```{r prep crayfish and agile frog dataset}
metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish") 
metrics_agile <- Pop_metrics %>% filter(Species == "Agile frog") 

metrics <- rbind(metrics_crayfish, metrics_agile)
metrics %<>% filter(Pond %in% c(1016, 111, 112, 66, 72, 888, 89, 91, 949))

metrics %>% group_by(TaxaUsed) %>% summarise(
                                             meanSEAc = mean(SEAc),
                                             meanNR = mean(NRcp),
                                             meanCDreg = mean(CDreg)
                                             )
```

### Wilcoxon tests
```{r wilcoxon tests agile frog crayfish}

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 1, p-value = 0.007812 **

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 5, p-value = 0.03906 *

wilcox.test(CDreg~TaxaUsed,paired=TRUE, data = metrics)
# V = 5, p-value = 0.03906 *
```

### Paired permutation tests 
```{r permut tests agile frog/crayfish}
metrics %<>% dplyr::select(Pond, NRcp, SEAc, CDreg, TaxaUsed)
metrics_pivoted <- pivot_wider(metrics, names_from = TaxaUsed, values_from = c(NRcp, SEAc, CDreg))

# Test par permutation sur la différence de moyenne :
# Calculate the difference between the two species
metrics_pivoted %<>% mutate(
                            diffNR = NRcp_Crayfish - `NRcp_Agile frog`,
                            diffSEAc = SEAc_Crayfish - `SEAc_Agile frog`,
                            diffCDreg = CDreg_Crayfish - `CDreg_Agile frog`
                            )


# set the number of permutations 
sim <- 10^4 #number of permutations

#create an empty vector to store the simulated averages
permNR <- rep(NA, sim)
permSEAc <- rep(NA, sim) 
permCDreg <- rep(NA, sim) 

k <- nrow(metrics_pivoted) # number of individuals to sample/permute
set.seed(2022) # make sure the results are going to be the same every time we run the code

# permute data
for (i in 1:sim) { #in each simulation
  # randomly select some pairs to have their difference multiplied by negative to indicate the other species should have the highest size
tirage = sample(c(-1,1), size = k, replace = T)
# append the average in the vector
permSEAc[i] = mean(metrics_pivoted$diffSEAc * tirage)
permNR[i] = mean(metrics_pivoted$diffNR * tirage)
permCDreg[i] = mean(metrics_pivoted$diffCDreg * tirage)
}

#add the observed difference as the final item in the vector
permNR[10000] = mean(metrics_pivoted$diffNR)
permSEAc[10000] = mean(metrics_pivoted$diffSEAc)
permCDreg[10000] = mean(metrics_pivoted$diffCDreg)

# Table disjonctive qui compte le nombre de valeurs dans perm qui sont superieures ou egales
# a la difference observee entre les moyennes, puis calcule la valeur de p

tablepropNR = table(abs(permNR) >= abs(mean(metrics_pivoted$diffNR)))
pNR = tablepropNR[2]/10000
pNR #p =0.0419

tablepropSEAc = table(abs(permSEAc) >= abs(mean(metrics_pivoted$diffSEAc)))
pSEAc = tablepropSEAc[2]/10000
pSEAc # p = 0.008

tablepropCDreg = table(abs(permCDreg) >= abs(mean(metrics_pivoted$diffCDreg)))
pCDreg = tablepropCDreg[2]/10000
pCDreg # p = 0.0342

```

\newpage
## B. Tree frog/All crayfish (n = 4)
### Prepare the dataset and show mean values
```{r test differences all crayfish and tree frog, echo = FALSE}
metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish")
metrics_treefrog <- Pop_metrics %>% filter(TaxaUsed == "Tree frog") 

metrics <- rbind(metrics_crayfish, metrics_treefrog)
metrics %<>% filter(Pond %in% c(66, 89, 91, 949))

metrics %>% group_by(TaxaUsed) %>% summarise(
                                             meanSEAc = mean(SEAc),
                                             meanNR = mean(NRcp),
                                             meanCDreg = mean(CDreg)
                                             )
```

### Wilcoxon tests
```{r wilcoxon tests tree frog crayfish, echo = FALSE}

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
#V = 2, p-value = 0.375

wilcox.test(CDreg~TaxaUsed,paired=TRUE, data = metrics)
#V = 4, p-value = 0.875

```


### Paired permutation tests
```{r permut tests tree frog crayfish, echo = FALSE}
metrics %<>% dplyr::select(Pond, NRcp, SEAc, CDreg, TaxaUsed)
metrics_pivoted <- pivot_wider(metrics, names_from = TaxaUsed, values_from = c(NRcp, SEAc, CDreg))


# Test par permutation sur la différence de moyenne: TAcp
# Calculate the difference
metrics_pivoted %<>% mutate(
                            diffNR = NRcp_Crayfish - `NRcp_Tree frog`,
                            diffSEAc = SEAc_Crayfish - `SEAc_Tree frog`,
                            diffCDreg = CDreg_Crayfish - `CDreg_Tree frog`
                            )


# permutation test: set the number of permutations 
sim <- 10^4 #number of permutations
permNR <- rep(NA, sim) #
permSEAc <- rep(NA, sim) #
permCDreg <- rep(NA, sim)

k <- nrow(metrics_pivoted)# number of individuals to sample/permute
set.seed(2022) # make sure the results are going to be the same

# permute data
for (i in 1:sim) {
  # randomly select some pairs to have their difference multiplied by negative to indicate the other species should have the highest size
tirage = sample(c(-1,1), size = k, replace = T)
permSEAc[i] = mean(metrics_pivoted$diffSEAc * tirage)
permNR[i] = mean(metrics_pivoted$diffNR * tirage)
permCDreg[i] = mean(metrics_pivoted$diffCDreg * tirage)
}

permNR[10000] = mean(metrics_pivoted$diffNR)
permSEAc[10000] = mean(metrics_pivoted$diffSEAc)
permCDreg[10000] = mean(metrics_pivoted$diffCDreg)

# Table disjonctive qui compte le nombre de valeurs dans perm qui sont superieures ou egales
# a la difference observee entre les moyennes
print("p-value for NR")
tablepropNR = table(abs(permNR) >= abs(mean(metrics_pivoted$diffNR)))
pNR = tablepropNR[2]/10000
pNR #p = 0.3816 no diff

print("p-value for SEAc")
tablepropSEAc = table(abs(permSEAc) >= abs(mean(metrics_pivoted$diffSEAc)))
pSEAc = tablepropSEAc[2]/10000
pSEAc 
tablepropSEAc
print("All simulations are TRUE, so p > 0.9999, no difference")

print("p-value for CDreg")
tablepropCDreg = table(abs(permCDreg) >= abs(mean(metrics_pivoted$diffCDreg)))
pCDreg = tablepropCDreg[2]/10000
pCDreg #p = 0.8803 no diff

```

\newpage
## C. Palmate newt/All crayfish (n = 5)
### Prepare the dataset and show mean values
```{r test differences crayfish adult and palmate newt, echo = FALSE}

metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish") 
metrics_palmate <- Pop_metrics %>% filter(TaxaUsed == "Palmate newt")

metrics <- rbind(metrics_crayfish, metrics_palmate)
metrics %<>% filter(Pond %in% c(111, 112, 66, 91, 949))

metrics %>% group_by(TaxaUsed) %>% summarise(
                                             meanSEAc = round(mean(SEAc),2),
                                             meanNR = round(mean(NRcp),2),
                                             meanCDreg = round(mean(CDreg),2)
                                             )
```

### Wilcoxon tests
```{r wilcoxon tests palmate newt crayfish, echo = FALSE}

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625 .

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 10, p-value = 0.625

wilcox.test(CDreg~TaxaUsed,paired=TRUE, data = metrics)
# V = 13, p-value = 0.1875

```


### Paired permutation tests 
```{r permut tests palmate newt crayfish, echo = FALSE}
metrics %<>% dplyr::select(Pond, NRcp, SEAc, CDreg, TaxaUsed)
metrics_pivoted <- pivot_wider(metrics, names_from = TaxaUsed, values_from = c(NRcp, SEAc, CDreg))

# Test par permutation sur la différence de moyenne: 
# Calculate the difference
metrics_pivoted %<>% mutate(
                            diffNR = NRcp_Crayfish - `NRcp_Palmate newt`,
                            diffSEAc = SEAc_Crayfish - `SEAc_Palmate newt`,
                            diffCDreg = CDreg_Crayfish - `CDreg_Palmate newt`
                            )


# permutation test: set the number of permutations 
sim <- 10^4 #number of permutations
permNR <- rep(NA, sim) #
permSEAc <- rep(NA, sim) #
permCDreg <- rep(NA, sim)

k <- nrow(metrics_pivoted)# number of individuals to sample/permute
set.seed(2022) # make sure the results are going to be the same

# permute data
for (i in 1:sim) {
  # randomly select some pairs to have their difference multiplied by negative to indicate the other species should have the highest size
tirage = sample(c(-1,1), size = k, replace = T)
permSEAc[i] = mean(metrics_pivoted$diffSEAc * tirage)
permNR[i] = mean(metrics_pivoted$diffNR * tirage)
permCDreg[i] = mean(metrics_pivoted$diffCDreg * tirage)
}

permNR[10000] = mean(metrics_pivoted$diffNR)
permSEAc[10000] = mean(metrics_pivoted$diffSEAc)
permCDreg[10000] = mean(metrics_pivoted$diffCDreg)


# Table disjonctive qui compte le nombre de valeurs dans perm qui sont superieures ou egales
# a la difference observee entre les moyennes
print("p-value for NR")
tablepropNR = table(abs(permNR) >= abs(mean(metrics_pivoted$diffNR)))
pNR = tablepropNR[2]/10000
pNR # p = 0.4966

print("p-value for SEAc")
tablepropSEAc = table(abs(permSEAc) >= abs(mean(metrics_pivoted$diffSEAc)))
pSEAc = tablepropSEAc[2]/10000
pSEAc # p = 0.0652

print("p-value for CDreg")
tablepropCDreg = table(abs(permCDreg) >= abs(mean(metrics_pivoted$diffCDreg)))
pCDreg = tablepropCDreg[2]/10000
pCDreg # p = 0.1935

```


\newpage
## D. Marbled newt/All crayfish (n = 0)
No stats


