---
title: "Stable coexistence between four native and one invasive species seen through the lens of isotopic niche analyses #6: Niche comparison"
author: 
- Nadege Belouard^[EcoBio, nadege.belouard@gmail.com]
- Jean-Marc Paillisson^[EcoBio, jean-marc.paillisson@univ-rennes1.fr]
date: "10/08/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2
  html_document:
    toc: TRUE
    toc_depth: 3
params:
  show_code: FALSE
  export_figures: TRUE
editor_options: 
  chunk_output_type: console
---

This vignette tests the link between the isotopic niche metrics of amphibians and candidate variables.


# 1. Setup

We begin by loading the required libraries and datasets.

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(dpi = 300, echo = params$show_code)

#Load packages
library(lme4)
library(car)
library(MuMIn)
library(ggplot2)
library(RVAideMemoire)
library(ggrepel)
library(magrittr)
library(Hmisc)
library(reshape2)
library(usdm)
library(tidyr)
library(dplyr)
library(tibble)
library(gridExtra)
library(here)
library(broom)
library(purrr)
```

```{r load datasets global}
Metrics_complete <- read.csv(file.path(here(), "exported_data", "Complete_metrics_species.csv"), h = T)
```


```{r load datasets per pop}
#Load dataset complete
Niche_metrics <- read.csv(file.path(here(), "exported_data", "Population_metrics_complete.csv"), h = T)
Mean_position <- read.csv(file.path(here(), "exported_data", "Meanposition_SEAc_complete.csv"), header = T)
names(Mean_position)[3] <- "Code"


#Load datasets without outliers
Niche_metrics_wo <- read.csv(file.path(here(), "exported_data", "Population_metrics.csv"), h = T)
Mean_position_wo <- read.csv(file.path(here(), "exported_data", "Meanposition_SEAc.csv"), header = T)
names(Mean_position_wo)[3] <- "Code"
```


```{r merge datasets}

# Merge niche metrics and mean position
# Pop_metrics <- merge(Niche_metrics_oneniche, Mean_position_oneniche, by = c("Pond", "Code"))
Pop_metrics <- merge(Niche_metrics, Mean_position, by = c("Pond", "Code"))
dim(Pop_metrics)[1] == dim(Niche_metrics)[1]
dim(Pop_metrics)[1] == dim(Mean_position)[1]

Pop_metrics_wo <- merge(Niche_metrics_wo, Mean_position_wo, by = c("Pond", "Code"))
dim(Pop_metrics_wo)[1] == dim(Niche_metrics_wo)[1]
dim(Pop_metrics_wo)[1] == dim(Mean_position_wo)[1]

# # Create the variable "Group" to differentiate newts and tadpoles
# for (i in 1:length(Pop_metrics$Pond)) {
#   if (Pop_metrics$Taxa[i] %in% c("Agile frog", "Tree frog")) {
#     Pop_metrics$Group[i] = "Tadpoles"
#     Pop_metrics$Clade[i] = "Amphibians"
#   } else if (Pop_metrics$Taxa[i] %in% c("Palmate newt", "Marbled newt")) {
#     Pop_metrics$Group[i] = "Newts"
#     Pop_metrics$Clade[i] = "Amphibians"
#   } else {
#     Pop_metrics$Group[i] = "Crayfish"
#     Pop_metrics$Clade[i] = "Crayfish"
#   }
# }

```


# 2. Compare crayfish and amphibians niches

## A. Check intra-group differences (DEPRECATED)
First we begin by checking whether we can group tadpoles together, newts together, and crayfish together

Crayfish adult/juvenile (n = 6) 
```{r test differences crayfish adult/juvenile}

# Composed estimates
crayfish_metrics <- Pop_metrics %>% filter(Pond %in% c(1016, 111, 112, 89, 91, 949), 
                                           Group == "Crayfish") %>% 
  group_by(Pond, Species) %>% 
  summarise(N = n(),
            TAcp = mean(TAcp),
            SEAccp = mean(SEAc),
            NRcp = mean(NRcp),
            CRcp = mean(CRcp),
            CDcp = mean(CDcp),
            MNNDcp = mean(MNNDcp),
            SDNNDcp = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg),
            TA = mean(TA),
            NR = mean(dY_range),
            CR = mean(dX_range),
            CD = mean(CD),
            MNND = mean(NND),
            SDNND = mean(SDNND)) %>% 
  pivot_longer(cols = c("TAcp", "SEAccp", "NRcp", "CRcp", "CDcp", "MNNDcp", "SDNNDcp", "D13C", "TP",
                        "TA", "NR", "CR", "CD", "MNND", "SDNND"), 
               names_to = "Metric", values_to = "Value") 


crayfish_metrics %>% 
  group_by(Metric) %>% nest() %>% 
  mutate(model = purrr::map(data, ~wilcox.test(Value ~ Species, data = .x, paired = TRUE))) %>% 
  mutate(tidy.model = purrr::map(model, ~broom::tidy(.x))) %>%   
  mutate(p.value = purrr::map_dbl(tidy.model, ~.x$p.value[1]))


# Boxplot
ggplot(aes(x = Metric, y = Value), data = crayfish_metrics) +
  geom_boxplot(aes(col = Species)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


# Difference between crayfish in NR for all estimates
# And difference in TA for the natural estimates only
```




## B. Check difference between crayfish and each amphibian species (DEPRECATED)

Agile frog/Crayfish adult (n = 9) 
```{r test differences crayfish adult and agile frog}
# Composed estimates
metrics <- Pop_metrics %>% filter(Pond %in% c(1016, 111, 112, 66, 72, 888, 89, 91, 949), 
                                           Species %in% c("Ad crayfish", "Agile frog")) %>% 
  group_by(Pond, Species) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


wilcox.test(TA~Species,paired=TRUE, data = metrics)
# V = 39, p-value = 0.05469 

wilcox.test(SEAc~Species,paired=TRUE, data = metrics)
# V = 39, p-value = 0.05469

wilcox.test(NR~Species,paired=TRUE, data = metrics)
# V = 37, p-value = 0.09766

wilcox.test(CR~Species,paired=TRUE, data = metrics)
# V = 32, p-value = 0.3008

wilcox.test(SDNND~Species,paired=TRUE, data = metrics)
# V = 36, p-value = 0.1289

wilcox.test(D13C~Species,paired=TRUE, data = metrics)
# V = 45, p-value = 0.003906 <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<

wilcox.test(TP~Species,paired=TRUE, data = metrics)
# V = 45, p-value = 0.003906 <<<<<<<<<<<<<<<<<<<<<<<<<<<<<

# Differ in D13C, TP
```
Tree frog/Crayfish adult (n = 4)
```{r test differences crayfish adult and tree frog}
# Composed estimates
metrics <- Pop_metrics %>% filter(Pond %in% c(66, 89, 91, 949), 
                                           Species %in% c("Ad crayfish", "Tree frog")) %>% 
  group_by(Pond, Species) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


wilcox.test(TA~Species,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(SEAc~Species,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(NR~Species,paired=TRUE, data = metrics)
# V = 2, p-value = 0.375

wilcox.test(CR~Species,paired=TRUE, data = metrics)
# V = 8, p-value = 0.375

wilcox.test(SDNND~Species,paired=TRUE, data = metrics)
# V = 5, p-value = 1

wilcox.test(D13C~Species,paired=TRUE, data = metrics)
#V = 10, p-value = 0.125

wilcox.test(TP~Species,paired=TRUE, data = metrics)
# V = 9, p-value = 0.25

# No difference found
```

Palmate newt/Crayfish adult (n = 5)
```{r test differences crayfish adult and palmate newt}
# Composed estimates
metrics <- Pop_metrics %>% filter(Pond %in% c(111, 112, 66, 91, 949), 
                                           Species %in% c("Ad crayfish", "Palmate newt")) %>% 
  group_by(Pond, Species) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


wilcox.test(TA~Species,paired=TRUE, data = metrics)
# V = 13, p-value = 0.1875

wilcox.test(SEAc~Species,paired=TRUE, data = metrics)
# V = 11, p-value = 0.4375

wilcox.test(NR~Species,paired=TRUE, data = metrics)
# V = 9, p-value = 0.8125

wilcox.test(CR~Species,paired=TRUE, data = metrics)
# V = 14, p-value = 0.125

wilcox.test(SDNND~Species,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625

wilcox.test(D13C~Species,paired=TRUE, data = metrics)
#V = 0, p-value = 0.0625

wilcox.test(TP~Species,paired=TRUE, data = metrics)
# V = 12, p-value = 0.3125

# No difference found
```

Marbled newt/Crayfish adult (n = 0)

Agile frog/Crayfish juv (n = 5)
```{r test differences crayfish juv and agile frog}
# Composed estimates
metrics <- Pop_metrics %>% filter(Pond %in% c(1016, 111, 112, 89, 91), 
                                           Species %in% c("Juv crayfish", "Agile frog")) %>% 
  group_by(Pond, Species) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


wilcox.test(TA~Species,paired=TRUE, data = metrics)
# V = 3, p-value = 0.3125

wilcox.test(SEAc~Species,paired=TRUE, data = metrics)
# V = 3, p-value = 0.3125

wilcox.test(NR~Species,paired=TRUE, data = metrics)
# V = 7, p-value = 1

wilcox.test(CR~Species,paired=TRUE, data = metrics)
# V = 4, p-value = 0.4375

wilcox.test(SDNND~Species,paired=TRUE, data = metrics)
# V = 5, p-value = 0.625

wilcox.test(D13C~Species,paired=TRUE, data = metrics)
# V = 0, p-value = 0.0625

wilcox.test(TP~Species,paired=TRUE, data = metrics)
# V = 0, p-value = 0.0625

# No difference found
```



Tree frog/Crayfish juv (n = 2)
Palmate newt/Crayfish juv (n = 4)
```{r test differences crayfish adult and palmate newt}
# Composed estimates
metrics <- Pop_metrics %>% filter(Pond %in% c(111, 112, 91, 949), 
                                           Species %in% c("Juv crayfish", "Palmate newt")) %>% 
  group_by(Pond, Species) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


wilcox.test(TA~Species,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(SEAc~Species,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(NR~Species,paired=TRUE, data = metrics)
# V = 2, p-value = 0.375

wilcox.test(CR~Species,paired=TRUE, data = metrics)
# V = 9, p-value = 0.25

wilcox.test(SDNND~Species,paired=TRUE, data = metrics)
# V = 6, p-value = 0.875

wilcox.test(D13C~Species,paired=TRUE, data = metrics)
#V = 0, p-value = 0.125

wilcox.test(TP~Species,paired=TRUE, data = metrics)
# V = 5, p-value = 1

# No difference found
```

Marbled newt/Crayfish juv (n = 0)


## C. Check differences between all crayfish and each amphibian species

### Complete dataset
Agile frog/All crayfish (n = 9)
```{r test differences all crayfish and agile frog}
# Composed estimates
metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish") 
metrics_agile <- Pop_metrics %>% filter(Species == "Agile frog") 

metrics <- rbind(metrics_crayfish, metrics_agile)
metrics$Pond
metrics %<>% filter(Pond %in% c(1016, 111, 112, 66, 72, 888, 89, 91, 949))

metrics %>% group_by(TaxaUsed) %>% summarise(medianTA = median(TAcp),
                                             medianSEAc = median(SEAc),
                                             medianNR = median(NRcp),
                                             medianCR = median(CRcp),
                                             # medianCD = median(CDcp),
                                             # medianMNND = median(MNNDcp),
                                             medianSDNND = median(SDNNDcp),
                                             medianD13C = median(D13Cavg),
                                             medianTP = median(TPavg))

wilcox.test(TAcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 1, p-value = 0.007812 **

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 1, p-value = 0.007812 **

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 5, p-value = 0.03906 *

wilcox.test(CRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 6, p-value = 0.05469 .

wilcox.test(SDNNDcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 7, p-value = 0.07422 .

wilcox.test(D13Cavg~TaxaUsed,paired=TRUE, data = metrics)
# V = 0, p-value = 0.003906 **

wilcox.test(TPavg~TaxaUsed,paired=TRUE, data = metrics)
# V = 0, p-value = 0.003906 **

# Difference found in TA, SEAc, NR, TP and D13C

# Replace wilcoxon tests by permutation tests
# Test
```

Tree frog/All crayfish (n = 4)
```{r test differences all crayfish and tree frog}
# Composed estimates
metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish")
metrics_treefrog <- Pop_metrics %>% filter(TaxaUsed == "Tree frog") 


metrics <- rbind(metrics_crayfish, metrics_treefrog)
metrics$Pond
metrics %<>% filter(Pond %in% c(66, 89, 91, 949))

wilcox.test(TAcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 6, p-value = 0.875

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
#V = 2, p-value = 0.375

wilcox.test(CRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 9, p-value = 0.25

wilcox.test(SDNNDcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 7, p-value = 0.625

wilcox.test(D13Cavg~TaxaUsed,paired=TRUE, data = metrics)
# V = 10, p-value = 0.125

wilcox.test(TPavg~TaxaUsed,paired=TRUE, data = metrics)
# V = 9, p-value = 0.25

# No difference
```

Palmate newt/All crayfish (n = 5)
```{r test differences crayfish adult and palmate newt}
# Composed estimates
metrics_crayfish <- Pop_metrics %>% filter(TaxaUsed == "Crayfish") 
metrics_palmate <- Pop_metrics %>% filter(TaxaUsed == "Palmate newt")

metrics <- rbind(metrics_crayfish, metrics_palmate)
metrics$Pond
metrics %<>% filter(Pond %in% c(111, 112, 66, 91, 949))

metrics %>% group_by(TaxaUsed) %>% summarise(medianTA = round(median(TAcp),2),
                                             medianSEAc = round(median(SEAc),2),
                                             medianNR = round(median(NRcp),2),
                                             medianCR = round(median(CRcp),2),
                                             # medianCD = round(median(CDcp),2),
                                             # medianMNND = round(median(MNNDcp),2),
                                             medianSDNND = round(median(SDNNDcp),2),
                                             medianD13C = round(median(D13Cavg),2),
                                             medianTP = round(median(TPavg),2))

wilcox.test(TAcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625 .

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625 .

wilcox.test(NRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 10, p-value = 0.625

wilcox.test(CRcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625 .

wilcox.test(SDNNDcp~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625 .

wilcox.test(D13Cavg~TaxaUsed,paired=TRUE, data = metrics)
# V = 0, p-value = 0.0625 .

wilcox.test(TPavg~TaxaUsed,paired=TRUE, data = metrics)
#V = 9, p-value = 0.8125

# No difference
```

Marbled newt/All crayfish (n = 0)



### Without outliers dataset
Agile frog/All crayfish (n = 9)
```{r test differences all crayfish and agile frog}
# Composed estimates
metrics_crayfish <- Pop_metrics_wo %>% filter(TaxaUsed == "Crayfish") %>% 
  group_by(Pond, TaxaUsed) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


metrics_agile <- Pop_metrics_wo %>% filter(Species == "Agile frog") %>% 
  group_by(Pond, TaxaUsed) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


metrics <- rbind(metrics_crayfish, metrics_agile)
metrics$Pond
metrics %<>% filter(Pond %in% c(1016, 111, 112, 66, 72, 888, 89, 91, 949))

wilcox.test(TA~TaxaUsed,paired=TRUE, data = metrics)
# V = 1, p-value = 0.007812 **

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 1, p-value = 0.007812 **

wilcox.test(NR~TaxaUsed,paired=TRUE, data = metrics)
# V = 5, p-value = 0.03906 *

wilcox.test(CR~TaxaUsed,paired=TRUE, data = metrics)
# V = 6, p-value = 0.05469 .

wilcox.test(SDNND~TaxaUsed,paired=TRUE, data = metrics)
# V = 7, p-value = 0.07422 .

wilcox.test(D13C~TaxaUsed,paired=TRUE, data = metrics)
# V = 0, p-value = 0.003906 **

wilcox.test(TP~TaxaUsed,paired=TRUE, data = metrics)
# V = 0, p-value = 0.003906 **

# Difference found in TA, SEAc, NR, TP and D13C
```

Tree frog/All crayfish (n = 4)
```{r test differences all crayfish and tree frog}
# Composed estimates
metrics_crayfish <- Pop_metrics_wo %>% filter(TaxaUsed == "Crayfish") %>% 
  group_by(Pond, TaxaUsed) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


metrics_treefrog <- Pop_metrics_wo %>% filter(TaxaUsed == "Tree frog") %>% 
  group_by(Pond, TaxaUsed) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


metrics <- rbind(metrics_crayfish, metrics_treefrog)
metrics$Pond
metrics %<>% filter(Pond %in% c(66, 89, 91, 949))

wilcox.test(TA~TaxaUsed,paired=TRUE, data = metrics)
# V = 6, p-value = 0.875

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(NR~TaxaUsed,paired=TRUE, data = metrics)
#V = 2, p-value = 0.375

wilcox.test(CR~TaxaUsed,paired=TRUE, data = metrics)
# V = 9, p-value = 0.25

wilcox.test(SDNND~TaxaUsed,paired=TRUE, data = metrics)
# V = 7, p-value = 0.625

wilcox.test(D13C~TaxaUsed,paired=TRUE, data = metrics)
# V = 10, p-value = 0.125

wilcox.test(TP~TaxaUsed,paired=TRUE, data = metrics)
# V = 9, p-value = 0.25

# No difference
```

Palmate newt/All crayfish (n = 5)
```{r test differences crayfish adult and palmate newt}
# Composed estimates
metrics_crayfish <- Pop_metrics_wo %>% filter(TaxaUsed == "Crayfish") %>% 
  group_by(Pond, TaxaUsed) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


metrics_palmate <- Pop_metrics_wo %>% filter(TaxaUsed == "Palmate newt") %>% 
  group_by(Pond, TaxaUsed) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 


metrics <- rbind(metrics_crayfish, metrics_palmate)
metrics$Pond
metrics %<>% filter(Pond %in% c(111, 112, 66, 91, 949))

wilcox.test(TA~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625

wilcox.test(SEAc~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625

wilcox.test(NR~TaxaUsed,paired=TRUE, data = metrics)
# V = 10, p-value = 0.625

wilcox.test(CR~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625

wilcox.test(SDNND~TaxaUsed,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625

wilcox.test(D13C~TaxaUsed,paired=TRUE, data = metrics)
# V = 0, p-value = 0.0625

wilcox.test(TP~TaxaUsed,paired=TRUE, data = metrics)
#V = 9, p-value = 0.8125

# No difference
```

Marbled newt/All crayfish (n = 0)

## D. Difference between crayfish (one niche) and amphibians (separate) (DEPRECATED)

Agile frog/Crayfish (n = 9)
```{r test differences all crayfish and agile frog}
# Composed estimates
metrics <- Pop_metrics %>% filter(Pond %in% c(1016, 111, 112, 66, 72, 888, 89, 91, 949), 
                                           Taxa %in% c("Crayfish", "Agile frog")) %>% 
  group_by(Pond, Taxa) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 

wilcox.test(TA~Taxa,paired=TRUE, data = metrics)
# V = 4, p-value = 0.02734 <<<<<<<<<<<<<<<<<<<<<<<<<<<

wilcox.test(SEAc~Taxa,paired=TRUE, data = metrics)
# V = 4, p-value = 0.02734 <<<<<<<<<<<<<<<<<<<<<<<<<<<

wilcox.test(NR~Taxa,paired=TRUE, data = metrics)
# V = 8, p-value = 0.09766

wilcox.test(CR~Taxa,paired=TRUE, data = metrics)
# V = 8, p-value = 0.09766

wilcox.test(SDNND~Taxa,paired=TRUE, data = metrics)
# V = 8, p-value = 0.09766

wilcox.test(D13C~Taxa,paired=TRUE, data = metrics)
# V = 0, p-value = 0.003906 <<<<<<<<<<<<<<<<<<<<<<<<<<<

wilcox.test(TP~Taxa,paired=TRUE, data = metrics)
# V = 0, p-value = 0.003906 <<<<<<<<<<<<<<<<<<<<<<<<<<<

# Difference found in TP and D13C
```
Tree frog/Crayfish (n = 4)
```{r test differences all crayfish and agile frog}
# Composed estimates
metrics <- Pop_metrics %>% filter(Pond %in% c(66, 89, 91, 949), 
                                           Taxa %in% c("Crayfish", "Tree frog")) %>% 
  group_by(Pond, Taxa) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 

wilcox.test(TA~Taxa,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(SEAc~Taxa,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(NR~Taxa,paired=TRUE, data = metrics)
# V = 1, p-value = 0.25

wilcox.test(CR~Taxa,paired=TRUE, data = metrics)
# V = 9, p-value = 0.25

wilcox.test(SDNND~Taxa,paired=TRUE, data = metrics)
# V = 4, p-value = 0.875

wilcox.test(D13C~Taxa,paired=TRUE, data = metrics)
# V = 10, p-value = 0.125

wilcox.test(TP~Taxa,paired=TRUE, data = metrics)
# V = 9, p-value = 0.25

# Difference found in TP and D13C
```
Palmate newt/Crayfish (n = 5)
```{r test differences all crayfish and agile frog}
# Composed estimates
metrics <- Pop_metrics %>% filter(Pond %in% c(111, 112, 66, 91, 949), 
                                           Taxa %in% c("Crayfish", "Palmate newt")) %>% 
  group_by(Pond, Taxa) %>% 
  summarise(N = n(),
            TA = mean(TAcp),
            SEAc = mean(SEAc),
            NR = mean(NRcp),
            CR = mean(CRcp),
            CD = mean(CDcp),
            MNND = mean(MNNDcp),
            SDNND = mean(SDNNDcp),
            D13C = mean(D13Cavg),
            TP = mean(TPavg)) 

wilcox.test(TA~Taxa,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625

wilcox.test(SEAc~Taxa,paired=TRUE, data = metrics)
# V = 14, p-value = 0.125

wilcox.test(NR~Taxa,paired=TRUE, data = metrics)
# V = 8, p-value = 1

wilcox.test(CR~Taxa,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625

wilcox.test(SDNND~Taxa,paired=TRUE, data = metrics)
# V = 15, p-value = 0.0625

wilcox.test(D13C~Taxa,paired=TRUE, data = metrics)
# V = 0, p-value = 0.0625

wilcox.test(TP~Taxa,paired=TRUE, data = metrics)
# V = 9, p-value = 0.8125

# Difference found in TP and D13C
```

